### 2020-05-25 paper/Manuscript.Rmd
Update paper

---
title: |
    | WORCS: A Workflow for Open Reproducible Code in Science
    | `r paste0("Version: ", worcs:::cran_version(), ", date: ", format(Sys.Date(), "%d %b,\ %Y"))`

shorttitle        : "WORKFLOW FOR OPEN SCIENCE"

author:
  - name: "Caspar J. Van Lissa"
    institution: "1,2"
    orcid: "0000-0002-0808-5024"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
  - name: "Andreas M. Brandmaier"
    institution: "3,4"
    orcid: "0000-0001-8765-6982"
  - name: "Loek Brinkman"
    institution: "2,5"
    orcid: "0000-0003-3997-1173"
  - name: "Aaron Peikert"
    institution: "3,6"
    orcid: "0000-0001-7813-818X"
  - name: "Barbara M.I. Vreede"
    institution: "7"
  - name: "Anna-Lena Lamprecht"
    institution: "2,8"
    orcid: "0000-0002-5023-4601"

affiliation:
  - id: "1"
    institution: "Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics"
  - id: "2"
    institution: "Open Science Community Utrecht"
  - id: "3"
    institution: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id: "4"
    institution: "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK"
  - id: "5"
    institution: "UMC Utrecht"
  - id: "6"
    institution: "Humboldt-Universitaet zu Berlin"
  - id: "7"
    institution: "University Library Utrecht"
  - id: "8"
    institution: "Utrecht University, Department of Information and Computing Sciences"

abstract: |
  Technological advances enable researchers to more easily conduct reliable,
  cumulative, collaborative science. Capitalizing on these advances has the
  potential to accelerate scientific progress. However, adopting open science
  principles can be challenging and costly, because doing so requires
  substantial conceptual education and training in the use of new tools. The 
  goal of this paper is to remove these obstacles. We introduce the Workflow for
  Open Reproducible Code in Science (WORCS). WORCS constitutes a step-by-step
  procedure that researchers can follow to make an entire project open and
  reproducible. It is based on best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies. To further facilitate the transition, the
  WORCS principles are implemented in an R-package, which offers an RStudio 
  project template that is, effectively, a "one-click solution" for R users. 
  This paper introduces the workflow, discusses how it meets different 
  requirements of open science, and provides examples of the implementation of
  `worcs` in R.

authornote: |
  This is a preprint, not a peer-reviewed publication. Comments, contributions,
  and coauthors are welcome. Please contact the lead author.
  Acknowledgements: Jeroen Ooms (feedback, suggesting gert, and using Remotes
  field).
keywords          : "open science; reproducibility; r; dynamic document generation; version control"
wordcount         : "5921"

bibliography      : ["worcs.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
classoption       : "noextraspace"
output            : papaja::apa6_pdf
urlcolor          : blue
knit              : worcs::cite_all
---

```{r load_packages, include = FALSE}
library("papaja")
mask <- TRUE
masked <- function(text){
  ifelse(mask, "[masked]", text)
}
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```

\setlength{\parskip}{0em}

<!--\vspace{\baselineskip}-->

The (social) sciences are arguably past the tipping point of a paradigm shift
towards open science.
Several immediate causes rallied support for this transition, including
highly publicized cases of scientific fraud [as discussed by
@leveltFailingScienceFraudulent2012], increasing awareness of questionable
research practices and their consequences for the credibility of findings [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
[@shroutPsychologyScienceKnowledge2018].
However, open science should not be seen as a cure (or punishment) for this
crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
[@tennantOpenScienceJust2018].
Moreover, open science creates opportunities for researchers to more easily
conduct reliable, cumulative, and 
collaborative science [@nosekScientificUtopiaOpening2012;
@adolphOpenBehavioralScience2012]. Open science also
promotes inclusivity, because it removes barriers for participation.
Capitalizing on these advances has the potential to accellerate scientific
progress [see also @coyneReplicationInitiativesWill2016].

Many individual researchers are motivated to adopt best practices for open
science and enjoy these benefits. And yet, the question of "how" to 
adopt open science practices can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by providing a simple workflow 
that meets most requirements open science, is based on best practices, and 
is supported by user-friendly software in R: The Worfkflow for Open Reproducible 
Code in Science (WORCS). WORCS is a lightweight approach to open science and
computational reproducibility. The workflow (Figure <!--\@ref(fig:workflow)-->) 
constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible. WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them. It can also be used in the absence of top-down support or guidelines
to produce work in accordance with best practices.
Although the principles underlying
the WORCS procedure are universal, they have been implemented for R users in the
R-package `worcs`, which offers an RStudio project template that is, effectively,
a "one click solution". This paper introduces the WORCS workflow, discusses how
it meets best practices for open science, and illustrates the use of the `worcs`
package.

<!-- AL: I second LB's comment somewhere below that the overall WORCS workflow should be introduced much earlier, maybe already here. A diagrammatic overview/flowchart of the principal steps would be great. It should catch the reader's eye and give an immediate idea of what WORCS is about. Now it is indeed quite buried in the details that follow. -->

<!--
thereby lowering the threshold for adoption.
, and still ensures reproducibility in most circumstances. 
 It is easy to implement, thereby lowering the threshold

* Grass-roots
* open science as opportunity
    - Easier work
    - Less mistakes
    - Better collaboration
    - Cumulative science
    - Error correction

Benefits to individual scientists: Automation (after a learning curve), thus more productive, avoid mistakes, join projects, clone projects and make them your own.
Collective benefits: Cumulative science, -->
<!-- AL: These "value propositions" would also be good to discuss already in the introduction. Directly after the introduction of the WORCS overview would make sense, I think. -->

<!-- AL: The remainder of the paper is structured as follows. ... Or some variation of this obligatory end-of-intro paragraph. As the paper touches on a lot of aspects, from principal to very practical, it would be very useful to provide some guidance here. 
CJ: I was always taught that you should not explain your structure in the paper - you should just structure the paper. If the structure is unclear, make it better. -->
<!--openness, transparency, and reproducible science -->

# Defining Open Science practices

Although Open Science is advocated by many, it does not have a unitary definition
[but, for an effort to ground open science in first principles, see 
@tennantValuePropositionOpen2020].
Instead, several guidelines exist for practices that are considered "open".
The TOP-guidelines are one of the most influential concrete operationalisations
of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis; and 8) replication of published results.
WORCS defines the goals of open science in terms of these guidelines, and is
designed to facilitate meeting each of these guidelines, with one exception: We
do not address replication of published results, because replication relates to
the subject of a research project, not to its execution.
We group the remaining seven guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

<!--## Goal of this paper

The TOP-guidelines are primarily geared towards top-down systemic change (what's in a name) by journals, funders, and organizations.
However, large organizations are slow to change.
For example, a snapshot of editorial practices indicated that, in 2015, only 3% out of 1151 psychology journals explicitly stated that
they accepted replication studies [@martinArePsychologyJournals2017; see also @klebelPeerReviewPreprint2020].
At the same time, <!-- edit -->
<!--The formation of "Open Science Communities" at universities around the globe
illustrates the momentum of this grassroots movement
[@brinkmanWidescaleAdoptionOpen2020]. -->

## Implementation in R

Although the principles underlying this workflow are universal, WORCS has first
been implemented for R users [@rcoreteamLanguageEnvironmentStatistical2020].
Several arguments support the choice to work in R.
First, R and all of it extensions are free open source software, which
should be the tool of choice for open science.
Second, as of this writing, R is the most advanced programming language in terms
of the implementation of tools required for an open science workflow.
Third, R is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fourth, R is well-supported; a vibrant online community exists for R, developing new R methods and packages, and providing support and tutorials for existing ones.
Finally, R is highly interoperable:
Packages are available to load nearly every imaginable filetype, and output can be written to most file types, including DOCX and PDF (in APA style), and HTML format. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from R, including C++, 
Fortran, and Python [@allaireMarkdownPythonEngine]. There are excellent free
resources for learning R, such as the book R for Data Science [@grolemundDataScience2017].

Working with R is simplified immensely by using the RStudio integrated
development engine (IDE) for R [@rstudioteamRStudioIntegratedDevelopment2015].
RStudio automates and streamlines tedious or complicated aspects of working with
the tools used in WORCS, and many are embedded directly into the visual user
interface.
This makes RStudio a comprehensive solution for open science research projects.
Another important feature of RStudio is project management.
A project bundles writing, analyses, data, references, etcetera, into a
self-contained folder, that can be uploaded entirely to 'GitHub' and downloaded
by future users. The `worcs` R-package installs a new RStudio project template.
When a new project is initialized from this template, the bookkeeping required
to set up an open science project is performed automatically. The `worcs` 
package has a [vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)
on how to set up your system with all of the required tools.

## Introducing the tools

### Dynamic document generation

WORCS relies on several free, open source software solutions.
The first is *dynamic document generation* (DDG): Writing scientific reports in
a format that interleaves written reports with blocks of code used to conduct
the analyses. The text is automatically formatted as a scientific paper in 
several potential styles. When the text is formatted, the results of the code
blocks are
automatically generated and insterted in the text, or rendered as Figures and
Tables. Dynamic document generation supercedes
the classical approach of using separate programs to write prose and conduct
analyses, and then manually copy-pasting analysis results into the text.

Although there is a slight learning curve to transitioning to DDG, the
investment really pays off. Time saved from painstakingly copy-pasting output
and manually formatting text soon outweighs the investment of switching to a new
program. Moreover, human error in manually copying results is eliminated.
When revisions require major changes to the analyses, all results, Figures and
Tables can be automatically updated. The flexibility in output formats also 
means that a manuscript can be rendered to presentation format, or even to a
website or blog post. Moreover, reproducibility is guaranteed, and can be
verified by reviewers and/or readers, because the code is run each time the
document is compiled. In sum, while writing academic papers in a programming environment might seem
counter-intuitive at first, this approach is much more amenable to the needs of
academics than most word processing software. It prevents mistakes, and saves
time.

In the R-implementation of WORCS, we recommend centering a research project
around one dynamically generated 'RMarkdown' document, which includes all 
analysis code. Longer scripts can be
stored in `.R` files, and called from the main document using the `source()`
function. When a reader or reviewer compiles this document, all code is run
automatically, thus verifying computational replicability. The 'RMarkdown' 
document can be automatically formatted in many styles, including
APA style [thanks to the R-package `papaja`, @austPapajaPrepareReproducible2020],
a host of other scientific formats 
[see @allaireRticlesArticleFormats2020], and as plain 'Markdown' for 'GitHub'.

### Version control

The second solution is *version control*:
Maintaining an indelible log of every change to all project files. Version 
control is a near-essential tool for scientific reproducibility, as anyone
learns who has had the experience of accidentally deleting a crucial file, or of
being unable to reproduce analyses because they ran some analyses interactively
instead of using syntax [see also @blischakQuickIntroductionVersion2016].
Many scientists use some form of *implicit* version control; for example, by
renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"),
tracking changes in word processing software, or using cloud hosting services
that retain backups of previous versions.

WORCS recommends using the *explicit* version control software 'Git'
(www.git-scm.com). Git tracks
changes to files, and stores these changes when the user makes a "commit" (a 
snapshot of the version controlled files).
Git retains a complete log of all commits, and users can compare changes between
different commits, or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's private computer.

The functionality of 'Git' is amplified by services such as 'GitHub'
(https://github.com). 'GitHub' is best understood as a cloud storage service
with social networking functionality. The cloud storage aspect
of 'GitHub' works as follows: You can "clone" (copy) a local 'Git' repository to
the 'GitHub' website, as a backup or research archive. The social network aspect
comes into play when a repository on 'GitHub' is set to "public": This allows 
other researchers to peruse the repository and see how the work was done, clone
it to their own computer to replicate the original work or apply the methods to
their own data, open "Issues" to ask questions or give feedback on the project,
or even send a "Pull request" with suggested changes to the text or code for
your consideration. 'Git' and 'GitHub' shine as tools for collaboration, because
different people can simultaneously work on different parts of a project, and
their changes can be compared and automatically merged on the website. Even on
solo projects, working with 'Git'/'GitHub' has many benefits: Staying organized,
being able to start a new study with a clone of an old, similar repository, or
splitting off an "experimental branch" to try something new, while retaining the
ability to "revert" (return) to a previous state of the project, or to "merge"
(incorporate) the experimental branch.

The R-implementation of WORCS uses the package `gert`, a 'Git' client for 'R'
[@oomsGertSimpleGit2019], to connect a local project to a remote repository on
'GitHub'. The `worcs` package also contains a user-friendly function to set
the default user credentials, `git_user()`, and a function to add, commit, and
push all changed files in one step, `git_update()`. It is worth noting that,
although this paper focuses on 'GitHub', which is the most widely used 'Git'
cloud storage service, two notable alternative services exist: 
'GitLab' (https://gitlab.com) and 'Bitbucket' (https://bitbucket.org). As of
version 0.1.2, the `worcs` template dialog accepts 'https' addresses of 
repositories on both of these platforms interchangeably. The package has a
[vignette](https://cjvanlissa.github.io/worcs/articles/git_cloud.html) with
additional instructions for using these platforms. However,
`worcs` has been most extensively tested with 'GitHub'.

### Dependency management

The third solution is *dependency management*: Keeping track of exactly what
software was used to conduct the analyses. At first glance, it might seem
sufficient to state that analyses were conducted in *Program X*. However, every
program is susceptible to changes, updates, and bugfixes. Open Source software,
in particular, is regularly updated because there is an
active community of developers contributing functionality and bugfixes. 
Potentially, any such update could change the results of the code, thus rendering
the analysis 'computationally non-reproducible'.

<!-- LB: and onwards: this section requires some background knowledge on R/packages/etc. This could be a place where readers stumble and stop reading. Perhaps introduce the concept of the R project template earlier, to see section on dependencies more into perspective. 
[main issue] it should become clear earlier what the WORCS workflow entails: e.g. switching to working with a particular R Project Template. Discuss the ‘workflow' section earlier. Related: Can you make a schematic overview of the WORCS?overview--> 
Many solutions exist to ensure computational reproducibility. These solutions typically work by
enveloping your research project in a distinct "environment" that only has access to programs that
are explicitly installed, and maintaining a record of these programs. These solutions differ in
user-friendliness and effectiveness. If strict reproducibility is required, users might choose to conduct all analyses in a 'Docker container': An environment that behaves like a
virtual computer, which can be stored like a sort of time capsule, and
identically reinstated on a user computer, or in the cloud.
This is effective, and also preserves software outside the R environment (e.g.,
preprocessing tools for neuroimaging data). An excellent Docker-based workflow
was recently introduced by @peikertReproducibleDataAnalysis2019. This workflow
is based on many of the same principles as WORCS, and the two approaches are
conceptually and practically compatible. However, the workflow described by 
Peikert and Brandmaier can be difficult to set up for novice users. An R-package
is currently in development to facilitate this transition [@repro]. Another 
user-friendly Docker-based workflow is offered by the cloud-based collaboration
platform 'Digital Ocean'. Although Digital Ocean enjoys some popularity among 
scientists, it is not clear whether this platform's
["Data Processing Agreement"](https://www.digitalocean.com/legal/data-processing-agreement/)
is GDPR compliant - it seems to focus more on limiting liability than on
compliance. The necessity to upload data for analysis is a risk with any
cloud-computing solution when human participants are involved.

For analyses conducted entirely in R, the recently released package `renv`
strikes a great balance between user-friendliness  and effectiveness. Developed by
the team behind 'RStudio', this package simply maintains a text-based,
human-readable log of all packages used, their version numbers, and where they
were installed from (e.g., 'CRAN', 'Bioconductor', 'GitHub'). As long as these
repositories maintain their archives, this solution is reliable and very
lightweight.
The text-based log file can be version controlled with 'Git'. `renv`
installs the logged packages in a cache that is shared between all
projects on the computer. Thus, if one of your projects uses a package that is
already in the cache, it does not
need to be installed again. This prevents long installation
times, large space requirements, and frustration having to re-install software
that is already installed on the system. When someone else loads your 
project, `renv` will install all of the required packages onto a cache on their
computer. 
This is not a foolproof method to ensure strict computational reproducibility, but it is more user-friendly
and lightweight than any alternative out there, requiring no technical skill to
setup. When choosing an appropriate solution for dependency management, it is
important to consider that all solutions have a limited shelf life
[@brownHowLearnedStop2017].
Ultimately, the best approach might be to use a "just good-enough" solution, and
acknowledge that all code requires some maintenance if you want to
reproduce it in the future. For projects in R, `renv` meets these requirements.

## Text-based files are better

A key consideration when developing a research project is what filetypes to use.
A case can be made to use text-based files exclusively (or as much as possible),
instead of binary files.
Text-based files can be read by machines and humans alike.
Binary files, such as 'Word' (`.docx`) or 'SPSS' (`.sav`, `.spo`) files,
must be decoded first, and many require commercial software to do so.
Text-based files are also required when using 'Git' for version control.
In 'Git', each change reflects a line of text added or removed. This means that, when
writing a version-controlled paper in plain-text, it is recommended to insert a
line break after every sentence (instead of after every paragraph). That way,
the change log will indicate which specific sentence was edited.
Binary files should not be version controlled with 'Git',
as changes in the encrypted file are uninterpretable. Binary files are also often 
larger than text-based files, which means they take up more space on cloud
hosting services, such as 'GitHub'. For these reasons, uploading large binary
files to these platforms is frowned upon.
'GitHub' encourages the use of text-based files by allowing visitors of their website to peruse
text-based files online, and
conveniently renders certain filetypes: For example, 'GitHub' displays `.md`
(Markdown) files as web pages, and `.csv` files as spreadsheets. This is why
`worcs` uses text-based file types like Markdown and `.csv` extensively.

# Introducing the workflow

## Preparing your system

Before you can use the R-implementation of the WORCS workflow, you have to install the required software.
You only have to perform this setup procedure only once for each computer you
intend to use `worcs` on. It should take about 30 minutes, and it is documented
in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
The setup vignette does not assume a prior installation of R, so it is suitable for novice users.
After setting up your system, you can use the WORCS workflow for all of your
projects. This step-by-step procedure is detailed in [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).
In this paper, we provide only a brief outline of the WORCS procedure (as of Version `r worcs:::cran_version()`) below.
We highly recommend users to refer to [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html) for the most up-to-date procedure (subject to improvements).

```{r results="asis", echo = FALSE, eval = TRUE}
get_workflow <- function(filename = "../vignettes/workflow.Rmd"){
  txt <- readLines(filename)
  txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
  txt <- gsub("(<!--S: |-->.*$)", "", txt)
  txt <- gsub("^#", "", txt)
  out <- vector("character")
  for(i in seq_along(txt)){
    if(startsWith(txt[i], "#")){
      out <- append(out, c("", txt[i], ""))
    } else {
      out <- append(out, txt[i])
    }
  }
  cat(out, sep = "\n")
}
get_workflow()
```

# How WORCS relates to the TOP-guidelines

## Comprehensive citation

The TOP-guidelines encourage comprehensive citation of literature, data, materials, methods,
and software. In principle,
researchers can meet this requirement by simply
citing every reference used. Unfortunately, citation of data and software is
less commonplace than citation of literature and materials. Crediting these
resources is important, because it incentivizes data sharing and the development
of open-source software, supports the open science efforts of others, and helps
researchers receive credit for *all* of their research output, in line with
[DORA](https://sfdora.org/).

To facilitate citing datasets, researchers sometimes publish *data papers*;
documents that detail the procedure, sample, and codebook. Specialized journals,
such as the
[*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/),
aid in the publication of these data papers. For smaller projects, researchers
sometimes publish their data in an online repository - like 'GitHub' - along
with a text file with the preferred citation for the data (which can be a
substantive paper), and the license that applies to the data, such as Creative
Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or
[BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt,
one can always contact the data creators and ask what the preferred citation is.

References for software are sometimes provided within the software environment;
for example, in R, package references can be found by
calling `citation("packagename")`. This returns an APA-style reference, and a
BibTeX entry. Software papers are also common, and are sometimes called
"Application Notes". The [Journal of Open Source Software (JOSS)](https://joss.theoj.org/)
offers programmers a way to generate a publication 
based on well-documented software;
in this process, the code itself is peer reviewed, and writing an actual paper is optional. For an example, see
@rosenbergTidyLPAPackageEasily2018.
Software citations can also sometimes be found 
in the source code repository (e.g., on 'GitHub'), using the
['Citation File Format'](https://citation-file-format.github.io/).

One important impediment to comprehensive citation is the fact that print journals operate with space constraints.
Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count.
Researchers can overcome this impediment by preparing two versions of the manuscript:
One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
The print version should reference the online version, so interested readers can find the comprehensive reference list.
The WORCS procedure suggests uploading the online version to a preprint server.
This is important because most preprint servers - at least those hosted by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
This means that authors will receive credit for cited work; even if they are cited only in the online version.
Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.

## Implementation in `worcs`

The `worcs` package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citing.html) to explain the
fundamentals for novice users, and offer recommendations specific to `worcs`. We
recommend using the free, open-source reference manager
[Zotero](https://www.zotero.org/download/); it is feature-rich, 
user-friendly, and highly interoperable with other reference managers. A 
tutorial for using Zotero with RMarkdown exists  [here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-rmarkdown/).

The `worcs` package also offers original solutions for comprehensive citation.
Firstly,
during the writing process, authors can mark the distinction between essential
and non-essential references. It is easier to do this from the start, instead of
going back to cut non-essential references just prior to publication. Standard
Markdown uses the at-symbol (`@`) to cite a reference. `worcs` additionally 
reserves the "double at"-symbol (`@@`) to cite a non-essential references.
Users can render the manuscript either with, or without, comprehensive citations
by adapting the `knit` command in the front matter, setting it to
`knit: worcs::cite_all` to render all citations, and to 
`knit: worcs::cite_essential` to remove all *non-essential* citations.

With regard to the citation of R-packages, it is worth noting that the `papaja` 
package [@austPapajaPrepareReproducible2020], which provides the 'APA6'
template used in `worcs`, also
includes functions to automatically cite all packages used in the R-session, and
add their references to the bibliography file.

## Data sharing

Data sharing is important for computational reproducibility and secondary
analysis. Computational reproducibility means that a third party can exactly
recreate the results from the original data, using the published analysis code.
Secondary analysis means that a third party can conduct sensitivity analyses,
explore alternative explanations, or even use existing data to answer a
different research question. From an open science perspective, data sharing is
always desirable. From a practical point of view, it is not always possible.
Data sharing may be impeded by lack of storage space, legal constraints, ethical restrictions, absence
of informed consent for data sharing, or privacy concerns. For example, one
concern that applies when working with human participant data, is
pseudonimization. 
Many legislatures require researchers to 'de-identify' human subject data upon collection, by removing or deleting any sensitive personal information and
contact details. However, when researchers have an imperfect understanding of the
legal obligations and dispensations, they can end up over- or undersharing, and
risk either placing participants at risk of being identified, or undermine the
replicability of their own work and potential reuse value of their data [@phillipsDiscombobulationDeidentification2016].

Once data are de-identified or pseudononimized, they can often be shared if participants have
provided informed consent to that effect. Nevertheles, it is important to note
that the European GDPR prohibits storing 'personal data' (information which can
identify a natural person whether directly or indirectly) on a server outside
the EU, unless it offers an "adequate level of protection". Although different
rules may apply to pseudonimized data, there are many repositories that are
GDPR compliant, such as the European servers of the Open Science Framework.
Different Universities, countries, and funding bodies also have their own
repositories that are complient with local legistation. Before sharing any
human participant data, it is recommended to obtain approval from an internal
(ethical) review board, guidance from Research Data Management Support, and
informed consent from participants.

<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data). E.g. can I indicate a point in my workflow (e.g. after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->
If data can be shared openly, they can simply be pushed to a public 'Git'
repository, along with the analysis code. This way, others can download the
entire repository and reproduce the analyses from start to finish. 
If data cannot be shared, researchers should aim to safeguard the potential for
computational reproducibility and secondary analysis as much as possible. WORCS
recommends two solutions to accomplish this goal.
The first solution is to publish a *checksum* of the original data file.
Think of a checksum as a 32-character summary, or as a "one word" description,
of the contents of a file.
Any change to the file will result in a different checksum. 
Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditer that running the public analysis code on the private data results in the published results of their work.

The second solution is to share a synthetic dataset with similar characteristics
to the real data. Synthetic data mimic the level of measurement and
(conditional) distributions of the real data
[see @nowokSynthpopBespokeCreation2016]. Sharing synthetic data allows any third
party to 1) verify that the published code works, 2) debug the code, and 3)
write valid code for alternative analyses. It is important to note that complex
multivariate relationships present in the real data are often lost in synthetic
data. Thus, findings from the real data might not be replicated in the synthetic
data, and findings in the synthetic data *should not be substantively interpreted*.
Still, sharing synthetic data facilitates data requests from third parties. A
third party can write analysis code based on the synthetic data, and send it to
the authors who evaluate it on the real data and send back the results. The
`worcs` package offers a simple but flexible function to generate synthetic data,
`synthetic()`, which by default uses random forests to generate a model for
synthesis (see @wrightRangerFastImplementation2015). For data synthesis to work
properly, it is essential that the `class` (data type) of variables is defined
correctly. As of `worcs` version `r worcs:::cran_version()`, numeric, integer, factor, and logical data
are supported out of the box. Other types of variables should be converted to
one of these types, or users can use a custom `model_expression` when calling
`synthetic`.

### Processing data in WORCS

Regardless of whether the data will be open or closed, it is important that the
raw data be left unchanged as much as possible. This eliminates human error.
Any alterations to the data - including processing steps and even error 
corrections - should be documented in the code, instead of applied to the raw data.
This way, the code will be a pipeline from the raw data to the published results.
Thus, for instance, if data have been entered manually in a spreadsheet, and 
a researcher discovers that in one of the variables, missing values were coded as
`-99` whereas they were coded as `-9` in all other variables, then we would recommend adding
a line of code to recode the missing values - thereby documenting the mistake - 
instead of "fixing" it by editing the raw spreadsheet.

When initializing a new `worcs` project, an empty R-script called `prepare_data.R` is generated.
As soon as raw data are collected, researchers should use this file to document
all teps necessary to load the 
data into R, pseudonimize it, and prepare it for analysis. 
For example, if the data was originally in SPSS format with IP addresses and GPS
location, this file might just contain the code required to read the SPSS file,
and to remove those columns of sensitive personal information. As soon as the
data are pseudonimized and processed into a 'tidy' format [i.e., data that can
be stored as a spreadsheet, see @wickhamTidyData2014], the researcher should
version control some indelible record of the raw data.

The `worcs` package offers two functions for version controlling a record of the
data: One for open, and one for closed data. Researchers should assume that the
decision to make data open is irreversible. Thus, the decision should be made well
before arriving at this stage in the workflow. If there is any doubt, it is prudent
to proceed with closed data and make the data open once all doubt is assuaged.

In `worcs`, researchers can call the function `open_data(data)` to make the data
publicly available. This function stores the `data` object (a `data.frame` or
`matrix`) in a spreadsheet file (called `data.csv` by default), and generates an
accompanying codebook that can be viewed online.
All changed files should now be (manually) added to the 'Git' repository, committed, and
pushed to GitHub. Once the GitHub repository is made public, these data will be
open to the public.
Assume that, once you do this, the data cannot be un-shared.

Alternatively, if the project requires data to remain closed, researchers can
call `closed_data(data)`. This function also stores the data in a local `.csv` file
and generates a codebook,
but the original data file is added to `.gitignore` so it cannot be
accidentally added to the 'Git' repository.
The function also computes a checksum for the original data, and logs it in the
`.worcs` project file. This checksum serves as the indelible record of the state 
of the raw data.
The function `closed_data()` also creates a synthetic dataset (called
`synthetic_data.csv` by default).
The user should push all changed files 'GitHub'.
Once the 'GitHub' repository is made public, people will have access to a
checksum for the original data that exist on your local machine, a codebook
describing those data, and a synthetic copy of the data.

<!-- Maybe remark that in the "closed" case, a the data should be backupped somewhere else, as GitHub will not do the job in this setting. -->

As the purpose of the `prepare_data.R` file is to prepare data for sharing in a
format as close to "raw data" as possible, the final line of this file
should typically be either
`open_data()` or `closed_data()`. 
The analysis code, on the other hand, should begin with `load_data()`.
This function loads the real data if it is available on the user's computer, and
otherwise, loads the synthetic data. This will have the effect that third party
users who copy a 'GitHub' repository with closed data will automatically load the synthetic
dataset, whereas the study authors, who have the real data stored locally, will
automatically load the real data. This function makes it possible for reviewers,
coauthors, and auditors, to write analysis scripts without requiring access to
the original data. They can simply start the script with `load_data()`, and
write their code based on the synthetic data. Then, they can submit their code
to you - by email or as a
["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request)
on GitHub. When you run their code on your system, `load_data()` will detect the
original data, and use that to run their script. You can then return the results
to the third party.

As of version `r worcs:::cran_version`, the `worcs` package only stores data as `.csv` files.
As explained above, this is recommended, because `.csv` files are human- and
machine readable, and because data typically need to be in a spreadsheet format
for analysis. Many types of data can be represented as a spreadsheet - text
corpora can have one row per document, and EEG- or EKG waveforms can have one
row per measurement occasion. The `prepare_data.R` file should document any
steps required to convert these data into tabular format. If it is data must be
stored in a different file type anyway, readers are encouraged to follow the
development of the `repro` package [@repro], which will enable such enhancements.

Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publically through GitHub. In this case, it is recommended to use `closed_data()`, and to manually upload the original `data.csv` file to the restricted access platform. If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the GitHub repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).

## Sharing code, research materials, design and analysis

When writing a manuscript created using dynamic document generation, analysis
code is embedded in the prose of the paper. Thus, the TOP-guideline of sharing
analysis code can be met simply by committing the source code of the manuscript
to the 'Git' repository, and making this repository Public, e.g., on 'GitHub'.
If authors additionally use open data and a reproducible environment (as suggested
in WORCS), then a third party can simply replicate all analyses by copying the
entire repository from the 'Git' hosting service, and Knitting the manuscript on
their local computer.

Aside from analysis code, the TOP-guidelines also encourage sharing new research
materials, and details of the study design and analysis. These goals can be
accomplished by placing any such documents in the 'Git' repository folder,
committing them, and pushing to a cloud hosting service. As with any document
version controlled in this way, it is advisable to use plain text only.

## Preregistration

Lindsay and colleagues [-@lindsayResearchPreregistration1012016] define
preregistration as "creating a permanent record of your study plans before you
look at the data. The plan is stored in a date-stamped, uneditable file in a
secure online archive." Two such archives are well-known in the social sciences:
[AsPredicted.org](https://aspredicted.org/), and [OSF.io](https://osf.io/).
However, 'Git' cloud hosting services also conform to these standards.
Thus, it is possible to preregister a study simply by pushing a text document
with the study plans to, e.g., 'GitHub'.
By tagging the commit as a preregistration
(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html)),
it is distinct from all other commits, and easily findable by people and programs.

The advantages, disadvantages, and pitfalls for preregistering different types
of studies have been extensively debated elsewhere [see @lindsayResearchPreregistration1012016].
For example, because the practice of preregistration has historically been closely
tied to experimental research,
it has been a matter of some debate whether secondary data
analyses can be preregistered 
[but see @westonRecommendationsIncreasingTransparency2019 for an excellent
discussion of the topic]. Preregistration is the only way to ensure that a
researcher is not "HARKing": Hypothesizing after the results are known
[@kerrHARKingHypothesizingResults1998].
When analyzing existing data, it is difficult to *prove* that a researcher did
not have direct (or indirect, through collaborators or by reading studies using
the same data) exposure to the data, before composing the preregistration.
However, the question of proof is only relevant from a perspective of preventing
scientific misconduct.
Good faith preregistration efforts always improve the quality of deductive
(theory-testing) research, because they avoid HARKing, ensure reliable
significance tests, avoid overfitting noise in the data, and limit the number of
forking paths researchers wander during data analysis
[@gelmanStatisticalCrisisScience2014].
In all cases, researchers should minimize exposure to the data, and disclose any
exposure.

WORCS takes the pragmatic position that, in deductive (hypothesis-testing) research,
it is beneficial to plan projects before executing them, to preregister these
plans, and adhere to them.
All deviations from this procedure should be disclosed; e.g., prior exposure to
the data, whether direct (e.g., by computing summary statistics) or indirect
(e.g., by reading papers using the same data); deviations from the analysis plan
to handle unforseen contingencies, such as violations of model assumptions;
or additional exploratory analyses.

WORCS recommends documenting a verbal, conceptual description of the study plans
in a text-based file. Optionally, an analysis script can be preregistered to
documenting the planned analyses.
If any changes must be made to the analysis code after obtaining the data,
one can refer to the conceptual description to justify the changes.
The ideal preregistered analysis script consists of a complete analysis
that can be evaluated once the data are obtained.
This ideal is often unattainable, because the data present researchers with
unanticipated challenges;
e.g., assumptions are violated, or analyses work differently than expected.
Some of these challenges can be avoided by simulating the data one expects to obtain,
and writing the analysis syntax based on the simulated data.
This topic is beyond the scope of the present paper, but many user-friendly
methods for simulating data are available in most statistical programming languages.
For instance, `R` users can use the package `simstudy` [@goldfeldSimstudySimulationStudy2020].

To preregister a project using 'Git', one can commit
all preregistration documents to the local 'Git' repository, 
and push them 
to a remote 'Git' hosting service.
By [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases) as "Preregistration" on the remote repository, it can be easily identified as
such and retrieved. This simple but robust approach renders pre-registration through
preregistration services, such as AsPredicted.org or OSF.io, optional.
However, it is easy to take this optional step when the preregistration is
written using dynamic document generation, as 
recommended in WORCS: This file can be formatted as PDF and uploaded as an attachment
to the formal preregistration service.

As soon as a project is preregistered on a Public 'Git' repository, it is
visible to the world, and reviewers (both formal reviewers designated by a
journal, and informal reviewers recruited by other means) can submit comments,
e.g., through pull requests on 'GitHub'. If the repository is Private, Reviewers
can be invited as "Collaborators". Private repositories can be made public at a
later date, along with their entire time-stamped history and tagged releases.
It is important to note that contributing to the repository in any way will void
reviewers' anonymity, as their contributions will be linked to a user name.

### Implementation of preregistration in `worcs`

Several excellent preregistration templates are available in the R-package
`prereg` [@austPreregMarkdownTemplates2019], including templates from
organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and
from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
When initializing an RStudio project with the `worcs` project template, one of
these preregistration templates can be selected, which will generate a file
called `preregistration.Rmd`. This file should be used to document study plans,
ideally prior to data collection.

## Compatibility with other standards for open science

Although WORCS was designed to meet the TOP-guidelines, it does so in a way that
is compatible with other open science principles. Most notably, WORCS is 
amenable to the
FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a]. These principles advocate that research
data should be Findable, Accessible, Interoperable and Reusable. They are
increasingly applied as a standard for other types of research output as well
[@lamprechtFAIRPrinciplesResearch2019].
WORCS helps users meet the FAIR principles by making all research objects -
including data and code - accessible in a repository that can be made public. 
The `worcs` package additionally
facilitates meeting the FAIR principles in several ways. First, each `worcs`
project has a YAML file (https://yaml.org/) that lists the research objects in
the project, which will allow web crawlers 
to index `worcs` projects. Second, a codebook is automatically generated for
each data file stored
using `open_data()` or `closed_data()`. We encourage users to elaborate on the
default codebook by adding
variable description and categories, which would - in the future - enable
indexing data by topic area. 
Incidentally, `worcs` is itself an example of FAIR research software [see @lamprechtFAIRPrinciplesResearch2019 and https://fair-software.eu/]. It is 
hosted on the public version controlled repository 'GitHub', it has an open
source licence (GPL v3.0), is registered in a community registry (CRAN), it
enables the citation of the software (using the `citation("worcs")` command, or
by citing this paper), and followed the CII Best Practices software quality
checklist during development.

Sharing all research objects, as advocated in WORCS, also provides a proper
basis for research evaluation
according to the San Francisco Declaration on Research Assessment (DORA;
https://sfdora.org/), which plays an increasing role in grant funding, hiring, 
and promotion procedures. Direct access to research objects allows stakeholders
to evaluate research quality based on content rather than relying on spurious
surrogate indicators like journal impact factors, conference rankings, and
h-indexes. The detailed version control and commit tracking of 'GitHub' and
similar platforms furthermore make it possible to assess the relative
contributions made by different researchers.

# Discussion

In this tutorial paper, we have presented a workflow for open reproducible code
in science.
The workflow aims to lower the threshold for grass-roots adoption of open science principles.
The workflow is supported by an R-package with an RStudio project template and 
convenience functions.
This relatively light-weight workflow meets most of the requirements for open science
as detailed in the TOP-guidelines, and is compatible with other open science guidelines.
The workflow helps researchers meet existing requirements for open science,
and to reap the benefits of working openly and reproducibly, 
even when such requirements are absent.

## Comparing WORCS to existing solutions

There have been several previous efforts to promote grass-roots adoption of open
science principles. Each of these efforts has a different scope, strengths, and
limitations that set it apart from WORCS. For example, there are "signalling
solutions"; guidelines to structure and incentivize disclosure about open
science practices. Specifically, @aalbersbergMakingScienceTransparent2018
suggested publishing a "TOP-statement" as supplemental material, which discloses
the authors' adherence to open science principles. Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based
Transparency Checklist that authors can complete online to generate a report.
Such signalling solutions are very easy to adopt, and they address
TOP-guidelines 1-7. Many journals now also offer authors the opportunity to earn
"badges" for adhering to open science guidelines
[@kidwellBadgesAcknowledgeOpen2016]. These signalling solutions help structure
authors' disclosures about, and incentivize adherence to, open science practices.
WORCS has its own checklist,

A different class of solutions instead focuses on the practical issue of *how*
researchers can meet the requirements of open science. <!--limitation of signalling approaches, that WORCS addresses, is that they do not clarify One could use WORCS to meet the TOP-guidelines, and document this in a TOP-statement.--> One notable example is the workflow for
reproducible analyses developed by @peikertReproducibleDataAnalysis2019. This 
excellent workflow strives to ensure strict computational reproducibility for
even the most sophisticated analyses. It has some limitations, however: In
focusing on reproducibility, however, it only addresses TOP-guidelines 2, 3, and
5. Moreover, in aiming for strict computational reproducibility it ends up being
relatively complex to implement. WORCS builds upon the same general principles
as Peikert and Brandmaier, which means the two workflows are compatible. What 
sets WORCS apart is that is an easy and sufficient solution for projects that
can be conducted entirely within R, resulting in a much more lightweight
approach to safeguard computational reproducibility under *most circumstances*.
WORCS also addresses a unique issue not covered by other existing solutions,
namely to provide a workflow most conducive to satisfying the TOP-guidelines and
FAIR principles, while being compatible with existing requirements.

```{r, echo = FALSE, eval = FALSE}
tab1 <- data.frame(
  Reference = c("Aalbers et al., 2018", "Peikert & Brandmaier, 2019"),
  Scope = c("Disclose adherence to open science principles", "Completely reproducible analysis"),
  Difficulty = c("Easy", "Hard"),
  Effectiveness = c("Low", "High"),
  "TOP-guidelines" = c("1-7", "2,3,5")
)
apa_table(tab1)
```

<!-- AL: Similar to my comment on the title, in this section it needs to make clear which scope of related work it covers. For example, in the life sciences there has also been a lot of work on practices for increasing reproducibility, with automated workflows, containerization, Research Objects, provenance protocols, and a lot more. I don't think it's necessary for this paper to discuss this, but that is under the provision that it defines its scope better. -->

## Limitations

WORCS is intended to substantially reduce the threshold for adopting best 
practices in open and reproducible research. However, several limitations and
issues for future development remain. One potential challenge is the learning
curve associated with the tools outlined in this paper. Learning to work with R,
Rmarkdown, and Git/GitHub requires an initial time investment. The amount of
time required is reduced by the availability of tutorials, such as this one, and
other tutorials cited throughout this document. Moreover, the time investment
tends to pay off. Working with R opens the door to many cutting edge analysis
techniques. Working with Rmarkdown saves time and prevents mistakes by avoiding
tedious copying of results into a text document. Working with Git/GitHub keeps
projects organized, prevents accidental loss of work, enables integrating
changes by collaborators in a non-destructive manner, and ensures that entire
research projects are archived and can be accessed or copied by third parties.
Thus, the time investment is eminently worthwhile.

Another important challenge is managing collaborations when only the lead author
uses Rmarkdown, and the coauthors use Word. In this case, it is possible to Knit
the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some
limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
When soliting feedback from co-authors, ask them to use Tracked Changes and
comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their
changes into the `manuscript.Rmd` file. In most cases, this is the most
user-friendly approach, as most lead authors would review changes by co-authors
anyway. A second approach is to ask collaborators to work in plain text. In this
case, send collaborators the `manuscript.Rmd` file, and ask them to open (and
save) it in Word or Notepad as a plain text file. When they send it back, make
sure any changes to your local file are committed, and then simply overwrite
your local version with their file. In RStudio, select the file in the Git tab,
and click the Diff button to examine what changes the collaborator has made
relative to your last committed version. Finally,
@peikertReproducibleDataAnalysis2019 mention the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
potentially useful way to collaborate with Word users. This package enables
bidirectional conversion between Word and
Rmarkdown, with support for tracked changes. However, `redoc` is currently in
suspended development, and for many users, a simple solution (i.e.,
asking collaborators to work in plain text) might be sufficient.

If all collaborators are committed to using WORCS, they can 
[Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo)
from the lead author on GitHub,
[clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository),
make their own changes, and
[send a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request)
to incorporate their changes. Working this way is extremely conducive to
scientific collaboration [@ramGitCanFacilitate2013]. Recall that, when using 
'Git' for collaborative writing, it is recommended to insert a
line break after every sentence so that the change log will indicate which
specific sentence was edited, and to prevent
["merge conflicts"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/resolving-a-merge-conflict-on-github)
when two authors edit the same line. The resulting document will be rendered to
PDF without spurious line breaks.

Moreover, being familiar
with GitHub opens doors to new forms of collaboration: In the open source
software community, continuous peer review and voluntary collaborative acts by
strangers who are interested in a project are commonplace
[see @adolphOpenBehavioralScience2012].  This kind of collaboration is budding
in scientific software development as well; for example, the lead author of this
paper became a co-author on several R-packages after submitting pull requests
with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
paper became involved by contributing pull requests to `worcs`. It is also
possible to invite such collaboration by 
[opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues)
for tasks that still need to be accomplished, and tag known Collaborators to
address them, or invite external collaborators to contribute their expertise.

## Future developments

The WORCS procedure and accompanying R-package provide a user-friendly and lightweight workflow for open, reproducible research in R, that meets all TOP-guidelines. It is beyond the scope of WORCS to support the incorporation of tools outside of the R-environment, or to containerize a project so that it can be identically reinstated on a different system or virtual machine. One important area of future development is to enable these extensions of the workflow. To this end, we are integrating WORCS into a modular framework of reproducibility tools for R [@repro]. This will make it possible for researchers to enhance a WORCS project with strict reproducibility tools like 'Docker' and 'Make', or vice versa, to initiate a WORCS project in a strictly reproducible environment. This combines the strengths of WORCS with those of @peikertReproducibleDataAnalysis2019.

<!-- Open peer review? -->

## Conclusion

WORCS offers a workflow for open reproducible code in science. The step-by-step procedure outlined in this tutorial helps researchers make an entire research project Open and Reproducible. The accompanying R-package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.

WORCS encourages and simplifies the adoption of Open Science principles in daily scientific work. It helps researchers make all research output created throughout the scientific process - not just manuscripts, but data, code, and methods - open and publicly assessible. This enables other researchers to reproduce results, facilitates cumulative science by allowing others to make direct use of these research objects.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\setlength{\parskip}{0em}

<div id = "refs"></div>
\endgroup