### 2020-10-08 inst/rstudio/templates/project/resources/pss.Rmd
Version 5

---
title           : "Title here"
shorttitle      : "Short title here"
date            : "`r Sys.setlocale('LC_TIME', 'C'); format(Sys.time(), '%d\\\\. %B %Y')`"
authors         : "Author names here"
---

<!-- PSS (Preregistration and Sharing Software) is a template that can be used
for pre-registrating confirmatory or exploratory studies.
For more details about the description please check the reference:

Krypotos, A. M., Klugkist, I., Mertens, G., & Engelhard, I. M. (2019).
A step-by-step guide on preregistration and effective data sharing for
psychopathology research. Journal of Abnormal Psychology, 128(6), 517-537.
https://psycnet.apa.org/doi/10.1037/abn0000424 -->

# Title

<!-- Here you can provide the project title. -->

## Authors

<!-- Here you can provide the authors' list. -->

## Affiliations

<!--  Authors' affiliations -->

# Background of the study

## Primary study/Secondary analyses

<!-- Is it a primary study (where data are collected) or a study 
where secondary analyses are performed on an existing data set? -->

## Does the study refer to a meta-analysis?

<!-- Mention whether the study refers to a meta-analysis.
In this case, you can ignore the irrelevant parts below
(e.g., number of participants). -->

## Research questions

<!-- Define the research question -->

## Study hypotheses

<!-- Define the study hypotheses -->

# Method

## Stimuli

<!-- Define the study stimuli -->

## Questionnaires

<!-- Define the study questionnaires, if used -->

## Equipment

<!-- Define the study equipment. This includes computer characteristics etc. -->

## Procedure

<!-- Define the study's procedure. -->

## Protocol

<!-- Define the study's protocol. -->

# Statistical analyses

## Participants

<!-- How many participants will be recruited? 
Has a power analysis been performed? -->

## Stopping rule

<!--  When will data accumulation end? -->

## Confirming hypotheses threshold

<!-- What is the threshold for confirming each of the research hypotheses? -->

## Disconfirming hypotheses threshold

<!-- What is the threeshold for rejecting each of the research hypotheses? -->

# Other

## Other (Optional)

<!-- Further comments -->

# References

<!-- Included references -->

## 
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
\noindent