### 2020-10-08 inst/rstudio/templates/project/resources/secondary.Rmd

---
title           : "Title here"
shorttitle      : "Short title here"
date            : "date here"
authors           : "authors here"
---

<!-- This template aims provides a relative short format to preregister
secondary analyses on preexisting data
(much like the AsPredicted.org format for new data collection).
For more details about the description please check the reference:

Mertens, G., & Krypotos, A. M. (2019). Preregistration of analyses of
preexisting data. Psychologica Belgica, 59(1), 338-352. doi: 10.5334/pb.493 -->

# Title

<!-- Here you can provide the project title.  -->

## Authors

<!-- Here you can provide the authors' list, together with the affiliations. -->

## Affiliations

<!--  Authors' affiliations-->

# Study hypotheses

<!-- Provide a brief description of the relevant theory 
and formulate the hypotheses as precisely as possible. -->

# Operationalization

<!-- State exactly how the variables specified in each hypothesis 
will be operationalized.  -->

# Data source

<!-- Specify the source of the obtained data. Also provide 
information about the context of the data source and clarify 
whether the data has been previously published. -->

# Data request/access

<!-- Specify how the data will be requested or accessed. 
Clarify whether the data were already available and whether 
the dataset has been previously explored or analyzed. -->

# Exclusion criteria

<!-- Specify whether there were any criteria for the 
exclusions of certain datasets, observations or time points.-->

# Statistical analyses

<!-- Specify the statistical model that will be used to analyze the data. 
Be as specific as possible and avoid ambiguity. -->

# Hypotheses (dis-)confirmation

<!-- Specify exactly how the hypothesis will be evaluated. 
Give specific criteria relevant to the used analytical 
model and framework (e.g., alpha-values, Bayes Factor, RMSEA).-->

# Analysis validation

<!-- Indicate whether the proposed analyses have previously been 
validated on a subset of the data, or a simulated dataset. 
If so, provide the data files and analysis code.-->

# Timeline

<!-- Provide the (foreseen) dates for the different steps 
in this preregistration form.-->

# What is known about the data that could be relevant for the tested hypotheses?

<!-- What is known about the data that could be relevant for the tested
hypotheses? I.e., disclose any prior exposure to the data set, direct or
indirect, and specific information (e.g., knowing the mean of a variable) that
is relevant for your research question. -->

## Other (Optional)

<!-- Further comments -->

# References

<!-- Included references -->

## 
\vspace{-2pc}
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{-1in}
\setlength{\parskip}{8pt}
\noindent