### 2020-11-14 paper/Manuscript_DS.Rmd
Better address FAIR principles

---
author:
  name: |
    [Caspar J. Van Lissa](c.j.vanlissa@uu.nl)$^{a,b}$
    Andreas M. Brandmaier$^{c,d}$
    Loek Brinkman$^{b,e}$
    Anna-Lena Lamprecht$^{b,f}$
    Aaron Peikert$^{c}$
    Marijn E. Struiksma$^{b,g}$
    Barbara Vreede$^{h}$

  affiliation: | 
      a: Utrecht University faculty of Social and Behavioral Sciences, department of Methodology & Statistics  
      b: Open Science Community Utrecht  
      c: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany  
      d: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany and London, UK  
      e: UMC Utrecht  
      f: Utrecht University, Department of Information and Computing Sciences  
      g: Utrecht University faculty of Humanities, department of Language, Literature & Communication, Utrecht Institute of Linguistics OTS  
      h: University Library Utrecht
title: "WORCS: A Workflow for Open Reproducible Code in Science"
short:      "WORKFLOW FOR OPEN SCIENCE"
shorttitle:      "WORKFLOW FOR OPEN SCIENCE"
abstract: |
  Adopting open science
  principles can be challenging and time-intensive, because doing so requires
  substantial conceptual education and training in the use of new tools.
  This paper introduces the Workflow for
  Open Reproducible Code in Science (WORCS): A step-by-step
  procedure that researchers can follow to make a research project open and
  reproducible.
  The purpose of the workflow is to lower the threshold for 
  adoption of open science principles.
  It is based on established best practices, and can be used either in
  parallel to, or in absence of, top-down requirements by journals,
  institutions, and funding bodies.
  To facilitate widespread adoption, the WORCS principles have been implemented
  in the R package `worcs`, which offers an RStudio 
  project template and utility functions for specific workflow steps.
  This paper introduces the conceptual workflow,
  discusses how it meets different 
  standards for open science, and addresses the functionality provided by the
  `R` implementation, `worcs`. 
  This paper is primarily targeted towards
  scholars conducting research projects in R,
  conducting research that involves academic prose, analysis code,
  and (optionally) tabular data.
  However, the workflow is flexible enough to accommodate other scenarios,
  and offers a sensible starting point for customized solutions.
  The source code for the `R` package and manuscript,
  and a list of [user examples of WORCS projects](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users), are available at 
  https://github.com/cjvanlissa/worcs.
authornote: "blabla"
keywords:
  # at least one keyword must be supplied
  formatted: [open science, reproducibility, r, dynamic document generation, version control]
  plain:     [open science, reproducibility, r, dynamic document generation, version control]
preamble: >
  \usepackage{amsmath}
bibliography      : ["worcs.bib"]  
urlcolor          : blue
output: rmarkdown::html_document
---
<!--
, "worcs_packages.bib"
output: 
  rmarkdown::html_document:
    citation_package: biblatex Note: Citations are always handled by pandoc when format is not PDF
knit: worcs::cite_all
csl: apa.csl
csl: apa.csl
-->

```{r load_packages, include = FALSE}
library("papaja")
mask <- TRUE
masked <- function(text){
  ifelse(mask, "[masked]", text)
}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message= FALSE)
```

# Introduction

Academia is arguably past the tipping point of a paradigm shift
towards open science.
Support for this transition was first motivated by several
highly publicized cases of scientific fraud [
@leveltFailingScienceFraudulent2012], increasing awareness of questionable
research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis
[@shroutPsychologyScienceKnowledge2018].
However, open science should not be seen as a cure (or punishment) for this
crisis. As the late dr. Jonathan Tennant put it: "Open science is just good science"
[@tennantOpenScienceJust2018].
Open science creates opportunities for researchers to more easily
conduct reliable, cumulative, and 
collaborative science [@nosekScientificUtopiaOpening2012;
@adolphOpenBehavioralScience2012]. Open science also
promotes inclusivity, because it removes barriers for participation.
Capitalizing on these advances has the potential to accelerate scientific
progress [see also @coyneReplicationInitiativesWill2016],
as has been aptly demonstrated by the role of open science in the response to the coronavirus pandemic [@sondervanCOVID19PandemicStresses2020].

Many researchers are motivated to adopt current best practices for open
science and enjoy these benefits. And yet, the question of how to 
get started can be daunting. Making the transition
requires researchers to become knowledgeable about different open science 
challenges and solutions, and to become proficient with new and unfamiliar tools.
This paper is designed to ease that transition by presenting a simple workflow,
based on established best practices,  
that meets most requirements open science:
The *Workflow for Open Reproducible Code in Science* (WORCS).

This paper introduces the conceptual workflow (referred to in upper case, "WORCS"),
and discusses its underlying principles and how it meets different standards for open science.
The principles underlying this workflow are universal and largely platform independent.
We also present a software implementation of WORCS for `R` users [@rcoreteamLanguageEnvironmentStatistical2020]:
The `R` package `worcs` [referred to in monospace font, @worcspackage2020].
This package offers a project template for RStudio
[@rstudioteamRStudioIntegratedDevelopment2015],
and several convenience functions to automate specific workflow steps.
This paper is most relevant for scholars using R to
conduct research that involves academic prose, analysis code,
and tabular data.
For other readers, it can serve as a primer on open science workflows,
and provide a sensible starting point for a customized solution.

The WORCS workflow constitutes a step-by-step procedure that researchers can follow to make a 
research project open and reproducible.
See Figure 1 for a schematic overview of the conceptual workflow,
which shows the steps to take in different stages of the research process.
The workflow is elaborated in greater detail below.
WORCS is compatible with open science
requirements already implemented by journals and institutions, and will help 
fulfill them.
It can also be used by individual authors to produce work in accordance with best practices
in the absence of top-down support or guidelines (i.e., "grass roots" open science).

```{r workflow, fig.cap="*Figure 1*: Schematic illustration of the WORCS procedure.", out.width="100%"}
knitr::include_graphics("workflow.png")
#knitr::include_graphics("workflow_graph/workflow.svg")
```

# Defining open science practices

Although open science is advocated by many,
it does not have a unitary definition.
Throughout this paper, we adhere to the definition of open science as
"the practice of science in such a way that others can collaborate and contribute,
where research data, lab notes and other research processes are freely available,
under terms that enable reuse, redistribution and reproduction of the research and its underlying data and methods" [@vicente-saezOpenScienceNow2018].
In this definition, the term "science" refers to all scholarly disciplines.
We further define the term *reproducible*, as used in the WORCS acronym, as being able to re-perform "the same analysis with the same code using a different analyst" [@patilVisualToolDefining2019].

The "TOP-guidelines" are one of the most influential concrete operationalisations 
of general open science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science,
the first seven of which "account for reproducibility of the reported results
based on the originating data, and for sharing sufficient information to conduct an independent 
replication" [@nosekPromotingOpenResearch2015a].
These guidelines are: 1) Comprehensive
citation of literature, data, materials, and methods; 2) sharing data, 3)
sharing the code required to reproduce analyses, 4) sharing new research
materials, and 5) sharing details of the design and analysis; 6)
pre-registration of studies before data collection, and 7) pre-registration of
the analysis plan prior to analysis.
The eighth criterion is "not formally a transparency standard for authors", but "addresses 
journal guidelines for consideration of independent replications for publication".
In this context, 
reproducibility can be defined as "re-performing the same analysis with the same code using a different analyst",
and replicability as "re-performing the experiment and collecting new data" [@patilVisualToolDefining2019].
WORCS defines the goals of open science in terms of the first seven TOP-guidelines,
and the workflow and its `R` implementation are designed to facilitate compliance therewith.
We group these guidelines into three categories: citation (1),
sharing (2-5), and preregistration (6-7).

## Introducing the tools

WORCS relies solely on a set of free, open source software solutions which we will
discuss before introducing the workflow.

### Dynamic document generation

The first is *dynamic document generation* (DDG): Writing scientific reports in
a format that interleaves written reports with blocks of code used to conduct
the analyses. The text is automatically formatted as a scientific paper in 
several potential styles. When the text is formatted, the code
blocks are
evaluated and their results are insterted in the text, or rendered as figures and
tables. Dynamic document generation supersedes
the classical approach of using separate programs to write prose and conduct
analyses, and then manually copy-pasting analysis results into the text.
This paper is an example of DDG, and its source code is
[available here](https://github.com/cjvanlissa/worcs/blob/master/paper/Manuscript_DS.Rmd).

Although transitioning to DDG involves a slight learning curve, we strongly
believe that the investment will pay off.
Time saved from painstakingly copy-pasting output
and manually formatting text soon outweighs the investment of switching to a new
program. Moreover, human error in manually copying results is eliminated.
When revisions require major changes to the analyses, all results, figures and
tables are automatically updated. The flexibility in output formats also 
means that a manuscript can be rendered to presentation format, or even to a
website or blog post. Moreover, the fact that code is evaluated each time the
document is compiled requires researchers to work reproducibly,
and allows reviewers and/or readers verify reproducibility simply by re-compiling
the document.
While writing academic papers in a programming environment might seem
counter-intuitive at first, this approach is much more amenable to the needs of
academics than most word processing software. It prevents mistakes, and saves
time.

In the `R` implementation of WORCS, we recommend centering a research project
around one dynamically generated `RMarkdown` document [@xieMarkdownDefinitiveGuide2018].
RMarkdown is based on [`Markdown`](https://en.wikipedia.org/wiki/Markdown),
which uses plain-text commands to typeset a document.
It enhances this format with support for R-code, and is compatible with [LaTeX](https://www.latex-project.org/) math typesetting.
For a full overview of the functionality of `RMarkdown`, see @xieMarkdownDefinitiveGuide2018.

The RMarkdown document can incorporate blocks of analysis code,
which can be used to render results directly to [Tables](https://bookdown.org/yihui/rmarkdown-cookbook/kable.html) and [Figures](http://zevross.com/blog/2017/06/19/tips-and-tricks-for-working-with-images-and-figures-in-r-markdown-documents/),
which can be [dynamically referenced](https://bookdown.org/yihui/rmarkdown-cookbook/cross-ref.html) in the text.
Long scripts can be stored in `.R` files,
and included in the main document using the [`read_chunk()`
function](https://github.com/yihui/knitr-examples/blob/master/113-externalization.Rmd) from the `knitr` package, which comes together with the `rmarkdown` package. When a reader or reviewer compiles this document, all code is run
automatically, thus verifying computational replicability. The `RMarkdown` 
document can be automatically formatted in many styles, including
APA style [thanks to the `R` package `papaja`, @austPapajaPrepareReproducible2020],
a host of other scientific formats 
[see @allaireRticlesArticleFormats2020], and as plain `Markdown`.

### Version control

The second solution is *version control*:
Maintaining an indelible log of every change to all project files. Version 
control is a near-essential tool for scientific reproducibility, as anyone
learns who has had the experience of accidentally deleting a crucial file, or of
being unable to reproduce analyses because they ran some analyses interactively,
instead of documenting the syntax in a script file [see also @blischakQuickIntroductionVersion2016].
Many scientists use some form of *implicit* version control; for example, by
renaming files after major changes (e.g., "manuscript_final_2.2-2019-10-12.doc"),
tracking changes in word processing software, or using cloud hosting services
that retain backups of previous versions.

An integral part of WORCS is the *explicit* version control software Git
(www.git-scm.com).
Git tracks changes to files on a line-by-line basis.
The user can store these changes by making a "commit": a 
snapshot of the version controlled files.
Each "commit" can contain as many changes as desired;
for example, a whole new paragraph, or many small changes made to address a single reviewer comment.
Each commit is further tagged with a message, describing the changes.
The `R` implementation of `worcs` contains a convenience function,
`git_update("your commit message")`,
which creates a new commit for all changes since the previous commit,
labels it with "your commit message", and pushes these changes to the remote repository.
It is, effectively, a Git "quick save" command.

From an open science perspective, Git is particularly appealing because it
retains a complete historical backlog of all commits.
This can be used, for example, to verify that authors executed the steps outlined in a preregistration.
Users can also compare changes between different commits,
or go back to a previous version of the code (for example,
after making a mistake, or to replicate a previous version of the results).
A project version controlled with Git is called a "Repository", or Repo.
Git only version controls files explicitly committed by the user. Moreover, it
is possible to prevent files from being version controlled - which is useful for
privacy sensitive data.
A `.gitignore` file lists files that should not be version-controlled.
These files thus exist only on the user's computer.

The functionality of Git is enhanced by services such as GitHub
(https://github.com). GitHub is best understood as a cloud storage service
with social networking functionality. The cloud storage aspect
of GitHub allows you to "clone" (copy) a local Git repository to
the GitHub website, as a backup or research archive. The social network aspect
comes into play when a repository on GitHub is set to "public": This allows 
other researchers to peruse the repository and see how the work was done; clone
it to their own computer to replicate the original work or apply the methods to
their own data; open "Issues" to ask questions or give feedback on the project,
or even send a "Pull request" with suggested changes to the text or code for
your consideration. Git and GitHub shine as tools for collaboration, because
different people can simultaneously work on different parts of a project, and
their changes can be compared and automatically merged on the website. Even on
solo projects, working with Git/GitHub has many benefits: Staying organized,
being able to start a new study with a clone of an old, similar repository, or
splitting off an "experimental branch" to try something new, while retaining the
ability to "revert" (return) to a previous state of the project, or to "merge"
(incorporate) the experimental branch.

Although GitHub is the most widely used
remote repository, `worcs` [supports alternative services](https://cjvanlissa.github.io/worcs/articles/git_cloud.html).
We will refer to all such services as "remote repositories".
The `worcs` package uses `gert` [@oomsGertSimpleGit2019] to connect a local project to a remote repository. It also includes user-friendly functions to set
user credentials, `git_user()`, and to add, commit, and
push changed files to a remote repository in one step, `git_update()`.

### Dependency management

The third solution is *dependency management*: Keeping track of exactly what
software was used to conduct the analyses. At first glance, it might seem
sufficient to state that analyses were conducted in *Program X*. However, every
program is susceptible to changes, updates, and bugfixes. Open Source software,
in particular, is regularly updated because there is an
active community of developers contributing functionality and bugfixes. 
Potentially, any such update could change the results of the code, thus rendering
the analysis computationally non-reproducible.

Many solutions exist to ensure computational reproducibility,
which differ in user-friendliness and effectiveness.
These solutions typically work by 
enveloping your research project in a distinct "environment" that only has access to programs that are explicitly installed, and maintaining a record of these programs.
When choosing an appropriate solution for dependency management, there is a tradeoff between ease-of-use on the one hand, and robustness on the other.
In making this tradeoff, it is important to consider that all solutions have a limited shelf life
[@brownHowLearnedStop2017].
Therefore, the best approach might be to use a "good-enough" solution, and
acknowledge that all code requires some maintenance if you want to
reproduce it in the future.

For dependency management, the `R` implementation of `worcs` relies on the recently released package `renv` [@usheyRenvProjectEnvironments2020].
This package strikes a great balance between user-friendliness and robustness.
Developed by the team behind RStudio, 
`renv` maintains a text-based,
human-readable log of all packages used, their version numbers, and where they
were installed from (e.g., CRAN, Bioconductor, GitHub).
This text-based log file can be version controlled with Git.
When someone else loads your project,
`renv` will install all of the required packages from this list.
This lightweight approach safeguards reproducibility as long as these repositories maintain their archives.
The `worcs` project template in RStudio automatically sets up `renv` for dependency management if the checkbox labeled "renv" is selected during project creation.

All solutions to computational reproducibility require users to make their dependencies explicit, often by reinstalling all software for each project in a controlled environment.
This can lead to long installation times and large memory requirements.
Similarly, `renv` requires users to call `install.packages()` for each library used - but it does not reinstall packages for each new project.
Instead, packages reside in a cache that is shared between all projects on the computer.
If a project uses a package that is already in the cache, it is not reinstalled, but linked from the cache.
This improves the user experience by saving time and memory.
As opposed to alternative solutions, `renv` is tightly integrated with the
native R package management system.
It monitors the scripts in the project directory for `library()` calls,
and notifies the user when it is necessary to update the dependencies file using the `snapshot()` function.
Therefore, `renv` does not require the users to change their workflow of installing/removing and using packages, and requires little technical skill to setup.
It is not a fail-safe method to ensure strict computational reproducibility,
but it meets most reproducibility requirements using only tools
native to `R` and RStudio.

When a project is conducted in a different software environment than `R`,
or if an R-project has external dependencies not tracked by `renv`,
then users might choose to use [Docker](https://www.docker.com/) for dependency management.  
Docker instantiates a "Docker container": An environment that behaves like a
virtual computer, which can be stored like a time capsule, and
identically reinstated on a user computer, or in the cloud.
Although the use of Docker falls outside of the scope of the present paper,
for existing users, a Docker build running `worcs` and its dependencies is available.
The [stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html)
explains how to instantiate this build.
For more information on this topic,
@peikertReproducibleDataAnalysis2019 describe 
a Docker-based workflow for computational reproducibility that is
conceptually and practically compatible with WORCS.

As Docker is somewhat challenging to set up for novice users,
an `R` package is currently in development to facilitate Docker integration [`repro`, @repro].
Another example of a Docker-based workflow is the cloud-based collaboration
platform 'Code Ocean'.
This platform greatly simplifies setting up a reproducible computing environment with version control and dependency management.
A key feature is that collaborators can log in to the server session,
rather than each having to set up Docker on a local computer.
One important consideration when using any cloud-computing solution
is the necessity to upload human participant data for analysis.
Such a platform would need to comply with relevant laws and ethical requirements.
For example, as of this writing, the [terms of use](https://codeocean.com/terms-of-use) for Code Ocean do not address the GDPR,
nor is it clear where data are stored.
This is a potential liability for European scholars.

## Text-based files are better

A key consideration when developing a research project is what filetypes to use.
WORCS encourages the use of text-based files whenever possible,
instead of binary files.
Text-based files can be read by machines and humans alike.
Binary files, such as Word (`.docx`) or SPSS (`.sav`, `.spo`) files,
must be decoded first, often requiring commercial software.
Although binary files can be version controlled with Git, 
their change log is uninterpretable.
Binary files are also often 
larger than text-based files, which means they take up more space on cloud
hosting services. For these reasons, uploading large binary
files to remote repositories is frowned upon.

Two additional points are worth noting:
First, as Git tracks line-by-line changes,
it is recommended to start each sentence on a new line.
That way, the change log will indicate which specific sentence was edited, 
instead of replacing an entire paragraph.
As RMarkdown does not parse a single line break as the end of a paragraph,
this practice does not disrupt the flow of a paragraph in the rendered version.
Second, the remote repository GitHub
renders certain text-based filetypes for online viewing: 
For example, `.md`
(Markdown) files are displayed as web pages, and `.csv` files as spreadsheets.
The `worcs` package uses this feature to, for example,
render a codebook for the data as `.md` so it will be shown
as a web page on GitHub.

# Introducing the workflow

We provide a conceptual outline of the WORCS procedure below, based on WORCS
Version 0.1.5.
The conceptual workflow recommends actions to take 
in the different phases of a research project, in order to work openly and transparently
in compliance with the TOP-guidelines and best practices for open science.
Note that, although the steps are numbered for reference purposes,
we acknowledge that the process of conducting research is not always linear.
We refer users of the `worcs` implementation in `R` to
[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/setup.html),
which describes in more detail which steps are automated by the `R` package
and which actions must be taken by the user.
The vignette is also updated with future developments of the package.

For an up-to-date list of sample `worcs` projects, see the [WORCS GitHub page](https://github.com/cjvanlissa/worcs).

```{r results="asis", echo = FALSE, eval = TRUE}
get_workflow <- function(filename = "../vignettes/workflow.Rmd"){
  txt <- readLines(filename)
  txt <- txt[grepl("<!--S:", txt, fixed = T)|grepl("^###", txt)]
  txt <- gsub("(<!--S: |-->.*$)", "", txt)
  txt <- gsub("^#", "", txt)
  out <- vector("character")
  for(i in seq_along(txt)){
    if(startsWith(txt[i], "#")){
      out <- append(out, c("", txt[i], ""))
    } else {
      out <- append(out, txt[i])
    }
  }
  cat(out, sep = "\n")
}
get_workflow()
```

## The R implementation of WORCS

WORCS is, first and foremost, a conceptual workflow that could be implemented in any software environment.
As of this writing, the workflow has been implemented for `R` users in the package `worcs` [@worcspackage2020].
Several arguments support our choice to implement this workflow first in `R` and RStudio,
although we encourage developers to port the software to other languages.
First, `R` and all of it extensions are free open source software, which
make it a tool of choice for open science.
Second, all tools required for an open science workflow are implemented in `R`,
most of these tools are directly accessible through the RStudio user interface,
and some of them are actively developed by the team behind RStudio.
Third, RStudio is a Public Benefit Corporation, whose commitment to open science is codified in the company charter:

> RStudio’s primary purpose is to create free and open-source software for data science, scientific research, and technical communication. This allows anyone with access to a computer to participate freely in a global economy that rewards data literacy; enhances the production and consumption of knowledge; and facilitates collaboration and reproducible research in science, education and industry.

Fourth, `R` is the second-most cited statistical software package [@muenchenPopularityDataScience2012], following SPSS, which has no support for any of the open science tools discussed in this paper.
Fifth, `R` is well-supported by
a vibrant and inclusive online community,
which develops new methods and packages,
and provides support and tutorials for existing ones.
Finally, `R` is highly interoperable:
Packages are available to read and write nearly every filetype,
including DOCX, PDF (in APA style), and HTML. 
Moreover, wrappers are available for many tools developed in other programming languages, and code written in other programming languages can be evaluated from `R`, including `C++`, 
`Fortran`, and `Python` [@allaireMarkdownPythonEngine2020]. There are excellent free
resources for learning `R` [e.g., @grolemundDataScience2017].

Working with `R` is simplified immensely by using the RStudio integrated
development engine (IDE) for `R` [@rstudioteamRStudioIntegratedDevelopment2015].
RStudio automates and streamlines tedious or complicated aspects of working with
the tools used in WORCS, and many are embedded directly into the visual user
interface.
This makes RStudio a comprehensive solution for open science research projects.
Another important feature of RStudio is project management.
A project bundles writing, analyses, data, references, et cetera, into a
self-contained folder, that can be uploaded entirely to a Git remote repository and downloaded
by future users. The `worcs` `R` package installs a new RStudio project template.
When a new project is initialized from this template, the bookkeeping required
to set up an open science project is performed automatically.

### Preparing your system

Before you can use the `R` implementation of the WORCS workflow,
you have to install the required software.
This 30 minute procedure is documented
in [the setup vignette](https://cjvanlissa.github.io/worcs/articles/setup.html).
After setting up your system, you can use `worcs` for all of your
projects.
Alternatively, Docker users can follow [the stetup with Docker vignette](https://cjvanlissa.github.io/worcs/articles/docker.html).
The `R` implementation of the workflow introduced earlier is detailed,
step-by-step, in
[the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html).

# How WORCS helps meet the TOP-guidelines

## Comprehensive citation

The TOP-guidelines encourage comprehensive citation of literature, data, materials, methods,
and software. In principle,
researchers can meet this requirement by simply
citing every reference used. Unfortunately, citation of data and software is
less commonplace than citation of literature and materials. Crediting these
resources is important, because it incentivizes data sharing and the development
of open-source software, supports the open science efforts of others, and helps
researchers receive credit for *all* of their research output, in line with the San Francisco
Declaration on Research Assessment, 
[DORA](https://sfdora.org/).

To facilitate citing datasets, researchers sometimes publish *data papers*;
documents that detail the procedure, sample, and codebook. Specialized journals,
such as the
[*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/),
aid in the publication of these data papers. For smaller projects, researchers often
simply share the data in a remote repository, along
with a text file with the preferred citation for the data (which can be a
substantive paper), and the license that applies to the data, such as Creative
Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or
[BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt,
one can always contact the data creators and ask what the preferred citation is.

References for software are sometimes provided within the software environment;
for example, in `R`, package references can be found by
calling `citation("packagename")`. This returns an APA-style reference, and a
BibTeX entry. Software papers are also common, and are sometimes called
"Application Notes". The [Journal of Open Source Software (JOSS)](https://joss.theoj.org/)
offers programmers a way to generate a publication 
based on well-documented software;
in this process, the code itself is peer reviewed, and writing an actual paper is optional. For an example, see
@rosenbergTidyLPAPackageEasily2018.
Software citations can also sometimes be found 
in the source code repository (e.g., on Git remote repositories), using the
['Citation File Format'](https://citation-file-format.github.io/). The
[*Software Heritage Project*](https://www.softwareheritage.org/2020/05/26/citing-software-with-style/)
recently published a BibLaTeX software citation style that will further facilitate
accurate citation of software in dynamically generated documents.

One important impediment to comprehensive citation is the fact that print journals operate with space constraints.
Print journals often discourage comprehensive citation, either actively, or passively by including the reference list in the manuscript word count.
Researchers can overcome this impediment by preparing two versions of the manuscript:
One version with comprehensive citations for online dissemination, and another version for print, with only the essential citations.
The print version should reference the online version, so interested readers can find the comprehensive reference list.
The WORCS procedure suggests uploading the online version to a preprint server.
This is important because most major preprint servers - including [arXiv.org](https://arxiv.org)
and all preprint services hosted by the [Open Science Framework (OSF)](https://help.osf.io/hc/en-us/articles/360019930493-Preprint-FAQs#how_do_i_find_preprints) - are indexed by Google Scholar.
This means that authors will receive credit for cited work; even if they are cited only in the online version.
Moreover, preprint servers ensure that the online version will have a persistent DOI, and will remain reliably accessible, just like the print version.

## Implementation in worcs

The `worcs` package includes a [vignette on citation](https://cjvanlissa.github.io/worcs/articles/citation.html) to explain the
fundamentals for novice users, and offer recommendations specific to `worcs`. We
recommend using the free, open-source reference manager
[Zotero](https://www.zotero.org/download/); it is feature-rich, 
user-friendly, and highly interoperable with other reference managers. A 
tutorial for using Zotero with RMarkdown exists  [here](https://christopherjunk.netlify.com/blog/2019/02/25/zotero-RMarkdown/).

The `worcs` package also offers original solutions for comprehensive citation.
Firstly,
during the writing process, authors can mark the distinction between essential
and non-essential references. It is easier to do this from the start, instead of
going back to cut non-essential references just prior to publication. Standard
Markdown uses the at-symbol (`@`) to cite a reference. `worcs` additionally 
reserves the "double at"-symbol (`@`) to cite a non-essential reference.
Users can render the manuscript either with, or without, comprehensive citations
by adapting the `knit` command in the front matter, setting it to
`knit: worcs::cite_all` to render all citations, and to 
`knit: worcs::cite_essential` to remove all *non-essential* citations.

With regard to the citation of `R` packages, it is worth noting that the `papaja` 
package [@austPapajaPrepareReproducible2020], which provides the `APA6`
template used in `worcs`, also
includes functions to automatically cite all packages used in the `R` session, and
add their references to the bibliography file.

## Data sharing

Data sharing is important for computational reproducibility and secondary
analysis. Computational reproducibility means that a third party can exactly
recreate the results from the original data, using the published analysis code.
Secondary analysis means that a third party can conduct sensitivity analyses,
explore alternative explanations, or even use existing data to answer a
different research question. 
If it is possible to share the data, these can simply be pushed to a Git remote
repository, along with documentation (such as a codebook) and the analysis code.
This way, others can download the
entire repository and reproduce the analyses from start to finish. 
From an open science perspective, data sharing is
always desirable. From a practical point of view, it is not always possible.

Data sharing may be impeded by several concerns.
One technical concern is whether data require special storage provisions,
as is often the case with "big data".
Most Git remote repositories do not support files larger than 100MB,
but it is possible to enhance a Git repository with [Git Large File Storage](https://git-lfs.github.com/) to add remotely stored files as large as several GB.
Files exceeding this size require custom solutions beyond the scope of this paper.

Sharing human participant data additionally entails legal and ethical concerns.
Before sharing any
human participant data, it is recommended to obtain approval from qualified
ethical and legal advisory organs, guidance from Research Data Management Support, and
informed consent from participants.
Many legislatures require researchers to 'de-identify' human subject data upon collection, by removing or deleting any sensitive personal information and
contact details. However, when researchers have an imperfect understanding of the
legal obligations and dispensations, they can end up over- or undersharing, and
risk either placing participants at risk of being identified, or undermine the
replicability of their own work and potential reuse value of their data [@phillipsDiscombobulationDeidentification2016].
Furthermore, it is important to note
that the European GDPR prohibits storing 'personal data' (information which can
identify a natural person whether directly or indirectly) on a server outside
the EU, unless it offers an "adequate level of protection". Although different
rules may apply to pseudonimized data, there are many repositories that are
GDPR compliant, such as the European servers of the Open Science Framework.
Different Universities, countries, and funding bodies also have their own
repositories that are complient with local legistation.
In sum, data should only be shared after consult qualified legal and ethical advisory organs,
obtaining informed consent from participants,
and de-identifying or pseudononimizing data.

If data cannot be shared, researchers should aim to safeguard the potential for
computational reproducibility and secondary analysis as much as possible. WORCS
recommends two solutions to accomplish this goal.
The first solution is to publish a *checksum* of the original data file [@rivest1992md5].
Think of a checksum as a 32-character unique identifier^[It is theoretically possible but improbable that random changes to a file will result in the same checksum.],
or as a lossy "summary" of the contents of a file.
Any change to the file will result in a different checksum. 
Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditor that running the public analysis code on the private data results in the published results of their work.

A second solution is to share a synthetic dataset with similar characteristics
to the real data. Synthetic data mimic the level of measurement and
(conditional) distributions of the real data
[see @nowokSynthpopBespokeCreation2016]. Sharing synthetic data allows any third
party to 1) verify that the published code works, 2) debug the code, and 3)
write valid code for alternative analyses. It is important to note that complex
multivariate relationships present in the real data are often lost in synthetic
data. Thus, findings from the real data might not be replicated in the synthetic
data, and findings in the synthetic data *should not be substantively interpreted*.
Still, sharing synthetic data facilitates data requests from third parties. A
third party can write analysis code based on the synthetic data, and send it to
the authors who evaluate it on the real data and send back the results. The
`worcs` package offers a simple but flexible function to generate synthetic data,
`synthetic()`, which by default uses random forests to generate a model for
synthesis (see @wrightRangerFastImplementation2015).
This function is based on the work by @nowokSynthpopBespokeCreation2016 (see the `R` package `synthpop`),
but with a simpler and more flexible user interface.
Specifically, the `worcs` function `synthetic()` accepts any function that can be used to model
marginal associations in the data.
For further details, see the [function documentation](https://cjvanlissa.github.io/worcs/reference/synthetic.html).

Note that for data synthesis to work
properly, it is essential that the `class` (data type) of variables is defined
correctly. As of `worcs` version `r worcs:::cran_version()`, numeric, integer, factor, and logical data
are supported out of the box. 
Other types of variables should be converted to one of these types.
It is important to consider this when using special data types,
such as the labeled vectors created by the `haven` package when reading an SPSS file.
We further caution that the default random forests algorithm is computationally intensive,
and may be unfeasible for use with "large" datasets.
Users can provide use a custom `model_expression` and `predict_expression`
to use a different algorithm when calling `synthetic()`.
For example, users could use `lm()` as a `model_expression` to use linear regression,
which preserves linear marginal relationships but can give rise to values out of range of the original data.
Or users could call `sample()` as a `predict_expression` to bootstrap each variable,
a very quick solution that maintains univariate distributions but loses all marginal relationships.

### Processing data in WORCS

When initializing a new `worcs` project, an empty `R` script called `prepare_data.R` is generated.
As soon as raw data are collected, researchers should use this file to document
all steps necessary to load the 
data into `R`, pseudonimize it, and prepare it for analysis.
As data should be shared in a format as close to raw as possible, 
this script should
be as short as possible and only document the minimum necessary steps.
For example, if the data was originally in SPSS format with IP addresses and GPS
location, this file might just contain the code required to read the SPSS file,
and to remove those columns of sensitive personal information. As soon as the
data are pseudonimized and processed into tabular (spreadsheet-like) format,
the researcher should
version control some indelible record of the raw data.
The concept of 'tidy' data
is a helpful standard; i.e., clean, tabular data that is formatted in a way that facilitates further analysis [see @wickhamTidyData2014].

One issue of concern is the practice of "fixing" mistakes in the data.
It is not uncommon for researchers to perform such corrections manually in the raw
data, 
even if all other analysis steps are documented.
Regardless of whether the data will be open or closed, it is important that the
raw data be left unchanged as much as possible. This eliminates human error.
Any alterations to the data - including processing steps and even error 
corrections - should be documented in the code, instead of applied to the raw data.
This way, the code will be a pipeline from the raw data to the published results.
Thus, for instance, if data have been entered manually in a spreadsheet, and 
a researcher discovers that in one of the variables, missing values were coded as
`-99` whereas they were coded as `-9` in all other variables, then we would recommend adding
a line of code to recode the missing values - thereby documenting the mistake - 
instead of "fixing" it by editing the raw spreadsheet.

The `worcs` package offers two functions for version controlling a record of the
data: One for open, and one for closed data. Researchers should assume that the
decision to make data open is irreversible. Thus, the decision should be made well
before arriving at this stage in the workflow. If there is any doubt, it is prudent
to proceed with closed data and make the data open once all doubt is assuaged.

In `worcs`, researchers can call the function `open_data(data)` to make the data
publicly available. This function stores the `data` object (such as a `data.frame` or
`matrix`) in a tabular data file.
It also generates an
accompanying codebook - 
a human-readable document that serves as a legend for the tabular data file -
that can be viewed online.
All changed files should now be added to the Git repository, committed, and
pushed to the remote repository.
This can be done through the Git panel in RStudio, or by running the `worcs`
function `git_update("commit message")`.
Once the remote repository is made public, these data will be
open to the public.
Assume that, once this is done, the data cannot be un-shared.

Alternatively, if the project requires data to remain closed, researchers can
call `closed_data(data)`. This function also stores the data in a local `.csv` file
and generates a codebook,
but the original data file is added to `.gitignore` so it cannot be
accidentally added to the Git repository.
The function also computes a checksum for the original data, and logs it in the
`.worcs` project file. This checksum serves as the indelible record of the state 
of the raw data.
The function `closed_data()` also calls the aforementioned `synthetic()` function to
create a synthetic dataset.
The user should push all changed files to the remote repository.
Once the remote repository is made public, people will have access to a
checksum for the original data that exist on your local machine, a codebook
describing those data, and a synthetic copy of the data.
Keep in mind that, when using closed data, the original data file exists only on
the user's computer. Adequate provisions to back up the data, while respecting
principles of responsible data stewardship, should be made.

As the purpose of the `prepare_data.R` file is to prepare data for sharing,
the final line of this file
should typically be either
`open_data()` or `closed_data()`.

After generating tidy, shareable data, this data is then loaded in the analysis code with the function `load_data()`.
This function only loads the real data if it is available on the user's computer, 
and otherwise, loads the synthetic data.
This will have the effect that third party
users who copy a remote repository with closed data will automatically load the synthetic
dataset, whereas the study authors, who have the real data stored locally, will
automatically load the real data.
This makes it possible for reviewers,
coauthors, and auditors, to write analysis scripts without requiring access to
the original data.
They can simply start the script with `load_data()`, and
write their code based on the synthetic data.
Then, they can submit their code
to you - by email or as a
["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request).
When you run their code on your system, `load_data()` will load the
original data, and use that to run their script. You can then return the results
to the third party.

Note that these functions can be used to store multiple data files, if necessary.
As of version 0.1.2, the `worcs` package only stores data as `.csv` files.
As explained above, this is recommended, because `.csv` files are human- and
machine readable, and because data typically need to be in tabular format
for analysis. Many types of data can be represented as a table; for example,
text corpora can have one row per document, and EEG- or ECG waveforms can have a
row per measurement occasion and a column per channel.
The `prepare_data.R` file should document any
steps required to convert these data into tabular format. 
If this is not an appropriate format for the data, readers are encouraged to follow the
development of the `repro` package [@repro], which will offer greater support for
reproducibly storing and retrieving diverse data formats.

Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publicly through a Git remote repository.
In this case, it is recommended to use `closed_data()`, and to manually upload the original `data.csv` file to the restricted access platform. If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the Git remote repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).

## Sharing code, research materials, design and analysis

When writing a manuscript created using dynamic document generation, analysis
code is embedded in the prose of the paper. Thus, the TOP-guideline of sharing
analysis code can be met simply by committing the source code of the manuscript
to the Git repository, and making this remote repository Public.
If authors additionally use open data and a reproducible environment (as suggested
in WORCS), then a third party can simply replicate all analyses by copying the
entire repository from the Git hosting service, and Knitting the manuscript on
their local computer.

Aside from analysis code, the TOP-guidelines also encourage sharing new research
materials, and details of the study design and analysis. These goals can be
accomplished by placing any such documents in the Git repository folder,
committing them, and pushing to a cloud hosting service. As with any document
version controlled in this way, it is advisable (but not required) to use plain text only.

## Preregistration

Lindsay and colleagues [-@lindsayResearchPreregistration1012016] define
preregistration as "creating a permanent record of your study plans before you
look at the data. The plan is stored in a date-stamped, uneditable file in a
secure online archive." Two such archives are well-known in the social sciences:
[AsPredicted.org](https://aspredicted.org/), and [OSF.io](https://osf.io/).
However, Git cloud hosting services also conform to these standards.
Thus, it is possible to preregister a study simply by committing a preregistration document
to the local Git repository, and pushing it to the remote repository.
Subsequently [taging the release](https://help.github.com/en/github/administering-a-repository/creating-releases)
as "Preregistration" on the remote repository renders it distinct from all other commits,
and easily findable by people and programs.
This approach is simple, quick, and robust.
Moreover, it is compatible with formal preregistration through services
such as AsPredicted.org or OSF.io:
If the preregistration is
written using dynamic document generation, as 
recommended in WORCS, this file can be rendered to PDF and uploaded as an attachment
to the formal preregistration service.

The advantages, disadvantages, and pitfalls for preregistering different types
of studies have been extensively debated elsewhere [see @lindsayResearchPreregistration1012016].
For example, because the practice of preregistration has historically been closely
tied to experimental research,
it has been a matter of some debate whether secondary data
analyses can be preregistered 
[but see @westonRecommendationsIncreasingTransparency2019 for an excellent
discussion of the topic].

When analyzing existing data, it is difficult to *prove* that a researcher did
not have direct (or indirect, through collaborators or by reading studies using
the same data) exposure to the data, before composing the preregistration.
However, the question of proof is only relevant from a perspective of preventing
scientific misconduct.
Preregistration is a good way to ensure that a
researcher is not "HARKing": Hypothesizing after the results are known
[@kerrHARKingHypothesizingResults1998].

Good faith preregistration efforts always improve the quality of deductive
(theory-testing) research, because they avoid HARKing, ensure reliable
significance tests, avoid overfitting noise in the data, and limit the number of
forking paths researchers wander during data analysis
[@gelmanStatisticalCrisisScience2014].

WORCS takes the pragmatic position that, in deductive (hypothesis-testing) research,
it is beneficial to plan projects before executing them, to preregister these
plans, and adhere to them.
All deviations from this procedure should be disclosed.
Researchers should minimize exposure to the data, and disclose any
prior exposure, whether direct (e.g., by computing summary statistics) or indirect
(e.g., by reading papers using the same data).
Similarly, one can disclose any deviations from the analysis plan
to handle unforeseen contingencies, such as violations of model assumptions;
or additional exploratory analyses.

WORCS recommends documenting a verbal, conceptual description of the study plans
in a text-based file. Optionally, an analysis script can be preregistered to
document the planned analyses.
If any changes must be made to the analysis code after obtaining the data,
one can refer to the conceptual description to justify the changes.
The ideal preregistered analysis script consists of a complete analysis
that can be evaluated once the data are obtained.
This ideal is often unattainable, because the data present researchers with
unanticipated challenges;
e.g., assumptions are violated, or analyses work differently than expected.
Some of these challenges can be avoided by simulating the data one expects to obtain,
and writing the analysis syntax based on the simulated data.
This topic is beyond the scope of the present paper, but many user-friendly
methods for simulating data are available in most statistical programming languages.
For instance, `R` users can use the package `simstudy` [@goldfeldSimstudySimulationStudy2020].

As soon as a project is preregistered on a Public Git repository, it is
visible to the world, and reviewers (both formal reviewers designated by a
journal, and informal reviewers recruited by other means) can submit comments,
e.g., through "Pull requests" on a Git remote repository. 
If the remote repository is Private, Reviewers
can be invited as "Collaborators". 
It is important to note that contributing to the repository in any way will void
reviewers' anonymity, as their contributions will be linked to a user name.
Private remote repositories can be made public at a
later date, along with their entire time-stamped history and tagged releases.

### Implementation of preregistration in worcs

The `worcs` workflow facilitates preregistration,
primarily by explicitly inviting a user to write the preregistration document.
To this end, `worcs` imports several preregistration templates from the `R` package
`prereg` [@austPreregMarkdownTemplates2019], including templates from
organizations like [AsPredicted.org](AsPredicted.org) and [OSF.io](OSF.io), and
from researchers [e.g., @vantveerPreregistrationSocialPsychology2016].
When initializing an RStudio project with the `worcs` project template, one of
these preregistration templates can be selected, which will generate a file
called `preregistration.Rmd`. This file should be used to document study plans,
ideally prior to data collection. 
Within the workflow, the Git cloud hosting services can be used to preregister a study simply by pushing a text document with the study plans to a Git remote repository.
By [tagging the commit](https://help.github.com/en/github/administering-a-repository/creating-releases) as a preregistration
(see [the workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html)),
it is distinct from all other commits, and easily findable by people and machines.

## Compatibility with other standards for open science

Aside from the TOP Guidelines, the FAIR Guiding Principles for scientific data management and stewardship [@wilkinsonFAIRGuidingPrinciples2016a] are an important standard for open science.
These principles advocate that digital research objects should be Findable, Accessible, Interoperable and Reusable.
Initially mainly promoted as principles for data, they are increasingly applied as a standard for other types of research output as well,
most notably software [@lamprechtFAIRPrinciplesResearch2019].

WORCS was not designed to comprehensively address these principles, but the workflow does facilitate meeting them.
In terms of Findability, all files in GitHub projects are searchable by content and meta-data, both on the platform and through standard search engines.
The workflow further recommends that users create a project page on the Open Science Framework – a recognized repository in the social sciences – and generate a DOI for this project.
Additional DOIs can be generated through the OSF for specific resources, such as data sets.
It is additionally or alternatively possible to connect the GitHub repository to Zenodo, which also offers DOIs and allows users to document projects with extensive metadata.
Finally, each worcs project has a [YAML file] (https://yaml.org/) that lists the data objects (and other meta-data) in the project, which will allow web crawlers to index worcs projects.
We use this metadata, for example, to identify public `worcs` projects on GitHub.
Accessibility is primarily safeguarded by hosting all project files in a public GitHub repository, which can be downloaded as a ZIP archive, and cloned or forked using the Git protocol. [GitHub is committed to long term accessibility of these resources]( https://archiveprogram.github.com/). We further encourage users to make their data accessible when possible.
When data must be closed, the `worcs` function `closed_data()` adds a paragraph to the project readme, indicating how readers can contact the author for access.

With regard to interoperability, the workflow recommends using plain-text files whenever possible, which ensures that code and meta-data are human- and machine-readable.
Furthermore, the `worcs` package is implemented in R, which is open source.
To promote reusability, we recommend users to specify an appropriate license for their projects.
The `worcs` package includes the CC licenses, and suggests a CC-BY 4.0 license by default.
This license is featured in the GitHub repository, and we encourage users to also display it on their project’s OSF page.
To further facilitate reusability, `worcs` automatically generates a codebook for
each data file.
We encourage users to elaborate on the default codebook by adding
variable description and categories, which would - in the future - enable
indexing data by topic area.

Incidentally, `worcs` itself follows the recommendations for [FAIR research software](https://fair-software.eu/). It is 
hosted on a version controlled public remote repository GitHub, has an open
source licence (GPL v3.0), is registered in a community registry (CRAN),
enables the citation of the software (using the `citation("worcs")` command, or
by citing this paper), and followed the [*CII Best Practices*](https://bestpractices.coreinfrastructure.org/en)
software quality checklist during development.

Sharing all research objects, as advocated in WORCS, also provides a thorough
basis for research evaluation
according to the San Francisco Declaration on Research Assessment (DORA;
https://sfdora.org/), which plays an increasing role in grant funding, hiring, 
and promotion procedures. Direct access to research objects allows stakeholders
to evaluate research quality based on content rather than relying on spurious
surrogate indicators like journal impact factors, conference rankings, and
h-indexes. The detailed version control and commit tracking of Git remote repositories
furthermore make it possible to assess the relative
contributions made by different researchers.

# Discussion

In this tutorial paper, we have presented a workflow for open reproducible code
in science.
The workflow aims to lower the threshold for grass-roots adoption of open science principles.
The workflow is supported by an `R` package with an RStudio project template and 
convenience functions.
This relatively light-weight workflow meets most of the requirements for open science
as detailed in the TOP-guidelines, and is compatible with other open science guidelines.
The workflow helps researchers meet existing requirements for open science,
and to reap the benefits of working openly and reproducibly, 
even when such requirements are absent.

## Comparing WORCS to existing solutions

There have been several previous efforts to promote grass-roots adoption of open
science principles. Each of these efforts has a different scope, strengths, and
limitations that set it apart from WORCS. For example, there are "signalling
solutions"; guidelines to structure and incentivize disclosure about open
science practices. Specifically, @aalbersbergMakingScienceTransparent2018
suggested publishing a "TOP-statement" as supplemental material, which discloses
the authors' adherence to open science principles. Relatedly, @aczelConsensusbasedTransparencyChecklist2019 developed a consensus-based
Transparency Checklist that authors can complete online to generate a report.
Such signalling solutions are very easy to adopt, and they address
TOP-guidelines 1-7. Many journals now also offer authors the opportunity to earn
"badges" for adhering to open science guidelines
[@kidwellBadgesAcknowledgeOpen2016]. These signalling solutions help structure
authors' disclosures about, and incentivize adherence to, open science practices.
WORCS, too, has its own checklist that calls attention to a number of concrete
items contributing to an open and reproducible research project. Users of the
`worcs` package can receive a badge, displayed on the `README.md` file,
based on a semi-automated scoring of the checklist items, by calling the function
`check_worcs()` within a `worcs` project directory.

A different class of solutions instead focuses on the practical issue of *how*
researchers can meet the requirements of open science.
A notable example is the workflow for reproducible analyses developed by @peikertReproducibleDataAnalysis2019,
which uses Docker to ensure strict computational reproducibility for
even the most sophisticated analyses.
There are some decisive differences with WORCS.
First, concerning the scope of the workflow: 
WORCS is designed to address a unique issue not covered by other existing solutions,
namely, to provide a workflow most conducive to satisfying TOP-guidelines 1-7 while adhering to the FAIR principles.
Peikert and Brandmaier instead focus primarily on computational reproducibility,
which is relevant for TOP-guidelines 2, 3, and 5.
Second, with regard to ease of use,
WORCS aims to bring down the learning curve by adopting sensible defaults for any decisions to be made.
Peikert and Brandmaier, by contrast, designed their workflow to be very flexible and comprehensive,
thus requiring substantially more background knowledge and technical sophistication from users.
To sum up, WORCS builds upon the same general principles
as Peikert and Brandmaier, and the two workflows are compatible. What sets WORCS apart is that it is more lightweight in terms of system and user requirements,
thereby facilitating adoption, but still ensures computational reproducibility under *most circumstances*.

One initiative that WORCS is fully compatible with, is the [*Scientific Paper of
the Future*](https://www.scientificpaperofthefuture.org/). This organization encourages
geoscientists to document data provenance and availability in public repositories,
document software used,
and document the analysis steps taken to derive the results. All of these goals
could be met using WORCS.

## Limitations

WORCS is intended to substantially reduce the threshold for adopting best 
practices in open and reproducible research. However, several limitations and
issues for future development remain. One potential challenge is the learning
curve associated with the tools outlined in this paper. Learning to work with `R`,
RMarkdown, and Git takes time - although tutorials such as this one substantially reduce the learning curve.
Moreover, the time investment
tends to pay off. Working with `R` opens the door to many cutting edge analysis
techniques. Working with RMarkdown saves time and prevents mistakes by avoiding
tedious copying of results into a text document. Working with Git keeps
projects organized, prevents accidental loss of work, enables integrating
changes by collaborators in a non-destructive manner, and ensures that entire
research projects are archived and can be accessed or copied by third parties.
Thus, the time investment is eminently worthwhile.

Another limitation is the fact that no single workflow can suit all research projects.
Not all projects conform to the steps outlined here;
some may skip steps, add steps, or follow similar steps in a different order.
In principle, WORCS does not enforce a linear order, and allows extentensive user flexibility.
A related concern is that some projects might require specialized solutions outside of the scope of this paper.
For example, some projects might use very large data files,
rely on data that reside on a protected server,
or use proprietary software dependencies that cannot be version controlled using `renv` or Docker.
In such cases, the workflow can serve as a starting point for a custom approach;
offering a primer on best practices, and a discussion of relevant factors to consider.
We urge users to inform us of such challenges and custom solutions,
so that we may add them to the [overview of WORCS-projects on GitHub](https://github.com/cjvanlissa/worcs), where they can serve as an example to others.

Another important challenge is managing collaborations when only the lead author
uses RMarkdown, and the coauthors use Word. In this case, it is possible to Knit
the manuscript to Word (.docx), by changing the line `output: papaja::apa6_pdf`
in the manuscript's YAML header to `output: papaja::apa6_docx`. There are some
limitations to the conversion, discussed [here](https://crsh.github.io/papaja_man/limitations.html#microsoft-word-documents).
When soliciting feedback from co-authors, ask them to use Track Changes and
comment bubbles in Word. <!-- LB: Overleaf?-->Then, manually incorporate their
changes into the `manuscript.Rmd` file. In most cases, this is the most
user-friendly approach, as most lead authors would review changes by co-authors
anyway. A second approach is to ask collaborators to work in plain text. In this
case, send collaborators the `manuscript.Rmd` file, and ask them to open (and
save) it in Word or Notepad as a plain text file. When they send it back, make
sure any changes to your local file are committed, and then simply overwrite
your local version with their file. In RStudio, select the file in the Git tab,
and click the Diff button to examine what changes the collaborator has made
relative to your last committed version. <!--Finally,
@peikertReproducibleDataAnalysis2019 mention the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@rossRedocReversibleReproducible2020] as a
potentially useful way to collaborate with Word users. This package enables
bidirectional conversion between Word and
RMarkdown, with support for tracked changes. However, `redoc` is currently in
suspended development, and for many users, a simple solution (i.e.,
asking collaborators to work in plain text) might be sufficient.-->

If all collaborators are committed to using WORCS, they can 
[Fork the repository](https://help.github.com/en/github/getting-started-with-github/fork-a-repo)
from the lead author on GitHub,
[clone it to their local device](https://help.github.com/en/github/creating-cloning-and-archiving-repositories/cloning-a-repository),
make their own changes, and
[send a pull request](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request)
to incorporate their changes. Working this way is extremely conducive to
scientific collaboration [@ramGitCanFacilitate2013]. Recall that, when using 
Git for collaborative writing, it is recommended to insert a
line break after every sentence so that the change log will indicate which
specific sentence was edited, and to prevent
["merge conflicts"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/resolving-a-merge-conflict-on-github)
when two authors edit the same line. The resulting document will be rendered to
PDF without spurious line breaks.

Being familiar
with Git remote repositories opens doors to new forms of collaboration:
In the open source
software community, continuous peer review and voluntary collaborative acts by
strangers who are interested in a project are commonplace
[see @adolphOpenBehavioralScience2012].  This kind of collaboration is budding
in scientific software development as well; for example, the lead author of this
paper became a co-author on several `R` packages after submitting pull requests
with bug fixes or additional functionality [@rosenbergTidyLPAPackageEasily2018; @hallquistMplusAutomationPackageFacilitating2018], and two co-authors of this
paper became involved by contributing pull requests to `worcs`. It is also
possible to invite such collaboration by 
[opening Issues](https://help.github.com/en/github/managing-your-work-on-github/about-issues)
for tasks that still need to be accomplished, and tag known collaborators to
address them, or invite external collaborators to contribute their expertise.

A final limitation is that, as WORCS is relatively new,
the number of current users is still limited.
We therefore do not have elaborate insights in user experiences or feedback.
We strive to make the uptake of worcs as easy as possible,
by means of this background paper, extensive documentation and vignettes for the `R` implementation,
[online video tutorials](https://www.youtube.com/playlist?list=PLOQIOFZl7uYbTgVdUee9n3eebMNmJtV9y),
webinars and [exemplar use cases](https://github.com/cjvanlissa/worcs/tree/master#adoption-of-worcs-by-users).
We encourage our users to provide us with feedback, feature requests, or bug reports through the channels [indicated here](https://github.com/cjvanlissa/worcs/tree/master#contributing-and-contact-information).

## Future developments

WORCS provides a user-friendly and
lightweight workflow for open, reproducible research, that meets all
TOP-guidelines pertaining to reproducibility (1-7).
Nevertheless, there are clear directions for future developments.
Firstly, although the workflow is currently implemented only in `R`, it is 
conceptually relevant for users of other statistical programming languages, and we welcome efforts to implement WORCS in other platforms.
We welcome efforts to implement WORCS in other platforms.
Secondly, even when a project is in `R`, it may have dependencies outside of the
`R` environment that cannot be managed using `renv`.
It is beyond the scope of the `worcs` package to support tools outside of `R`,
or to containerize a project so that it can
be identically reinstated on a different system or virtual machine.
The forthcoming package `repro` [@repro] will offer such extensions of the
workflow, thus combining the strengths of WORCS with the solutions proposed
by @peikertReproducibleDataAnalysis2019.

<!-- Open peer review? -->

## Conclusion

WORCS offers a workflow for open reproducible code in science. The step-by-step procedure outlined in this tutorial helps researchers make an entire research project Open and Reproducible. The accompanying `R` package provides user-friendly support functions for several steps in the workflow, and an RStudio project template to get the project set up correctly.

WORCS encourages and simplifies the adoption of open science principles in daily scientific work. It helps researchers make all research output created throughout the scientific process - not just manuscripts, but data, code, and methods - open and publicly assessible. This enables other researchers to reproduce results, and facilitates cumulative science by allowing others to make direct use of these research objects.

# References