### 2021-02-03 README.Rmd
try to add github actions CI

---
output: github_document
---

# WORCS <a href='https://osf.io/zcvbs/'><img src='https://github.com/cjvanlissa/worcs/raw/master/docs/worcs_icon.png' align="right" height="139" /></a>

<!-- README.md is generated from README.Rmd. Please edit that file -->
[![CRAN status](https://www.r-pkg.org/badges/version/worcs)](https://cran.r-project.org/package=worcs)
[![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/grand-total/worcs?color=blue)](https://r-pkg.org/pkg/worcs)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![check-standard](https://github.com/cjvanlissa/worcs/workflows/check-standard/badge.svg)](https://github.com/cjvanlissa/worcs/actions)
[![CII Best Practices](https://bestpractices.coreinfrastructure.org/projects/3969/badge)](https://bestpractices.coreinfrastructure.org/projects/3969)
<!--[![Codecov test
coverage](https://codecov.io/gh/cjvanlissa/worcs/branch/master/graph/badge.svg)](https://codecov.io/gh/cjvanlissa/worcs?branch=master)-->
<!--[![DOI](http://joss.theoj.org/papers/10.21105/joss.00978/status.svg)](https://doi.org/10.21105/joss.00978)-->

The Workflow for Open Reproducible Code in Science (WORCS) is an easy to adopt approach to ensuring a research project meets the requirements of Open Science from the start. It is based on a "good enough" philosophy, prioritizing user-friendliness over exhaustiveness. It can be used either in absence of, or in parallel to, existing requirements for Open workflows. It can also be enhanced with more elaborate solutions for specific issues.

## Where do I start?

For most users, the recommended starting point is to [read the paper](https://osf.io/zcvbs/), currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press),
which introduces the WORCS workflow, explains the underlying tools, and illustrates how the `worcs` package can be used to create a new project that follows the workflow.

The workflow is illustrated below; the [workflow vignette](https://cjvanlissa.github.io/worcs/articles/workflow.html) describes each step in detail.

```{r, echo = FALSE}
knitr::include_graphics("https://github.com/cjvanlissa/worcs/raw/master/paper/workflow_graph/workflow.png")
```

## Installing the package

Before installing the package, please read [this vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), which explains how to set up your computer for `worcs`.

After reading [the vignette](https://cjvanlissa.github.io/worcs/articles/setup.html), you can install the development version of the `worcs` package from GitHub with:

``` r
if(!requireNamespace("remotes"))install.packages("remotes")
remotes::install_github("cjvanlissa/worcs", dependencies = TRUE, update = "never")
tinytex::install_tinytex()
worcs::git_user("your_name", "your_email")
```

If you intend to write APA style manuscripts, you can additionally install the `papaja` package:

``` r
remotes::install_github("crsh/papaja", dependencies = TRUE, update = "never")

```

## Citing WORCS

You can cite WORCS using the following citation (please use the same citation for either the package, or the paper):

> Van Lissa, C. J., Brandmaier, A. M., Brinkman, L., Lamprecht, A., Peikert, A., , Struiksma, M. E., & Vreede, B. (in press). WORCS: A Workflow for Open Reproducible Code in Science. Data Science, in press. https://doi.org/10.17605/OSF.IO/ZCVBS

## About this repository

This repository contains the following:

1.  An R-package called `worcs`, with convenience functions to facilitate the WORCS workflow.
2.  In the subfolder `./paper`, the source files for the paper describing the WORCS workflow.

The repository serves two functions: To allow users to install the `worcs` package, and to allow collaborators access to the source code for the package and paper.

## Repository structure

```{r echo = FALSE, results = "asis"}
descripts <- data.frame(matrix(c(
  "_pkgdown.yml", "YAML for package website", "do not edit",
  "DESCRIPTION", "R-package DESCRIPTION", "do not edit",
  "LICENSE.md", "Project license", "do not edit",
  "NAMESPACE", "R-package namespace", "machine-written",
  "README.md", "Read this file to get started! ", "do not edit",
  "README.Rmd", "R-markdown source for readme.md", "human editable",
  "worcs.Rproj", "RStudio project file", "do not edit",
  "docs/", "Package website", "machine-written",
  "inst/", "RStudio project template files ", "human editable",
  "man/", "R-package documentation", "do not edit",
  "paper/", "WORCS paper source files", "human editable",
  "R/", "R-package source code", "human editable",
  "vignettes/", "R-package vignettes", "human editable"
), ncol = 3, byrow = TRUE))
names(descripts) <- c("File", "Description", "Usage")
knitr::kable(descripts)
```

## Adoption of WORCS by users

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(gh)
library(dlstats)
dl <- dlstats::cran_stats("worcs")
dl <- sum(dl$downloads)

worcs_repo <- gh("/repos/cjvanlissa/worcs")

repos <- gh("GET /search/code", username = "cjvanlissa", q="worcs_version filename:.worcs")

worcs_repos <- data.frame(t(sapply(repos[["items"]], function(x){
  c(owner = x$repository$owner$login,
    name = x$repository$name,
    url = x$repository$html_url
  )
})))

worcs_repos <- worcs_repos[!((worcs_repos$owner %in% c("cjvanlissa", "cran") & worcs_repos$name == "worcs")| grepl("(penguin|\\btmp|demo|test)", tolower(worcs_repos$name))), ]

tmp <- paste0(sapply(unique(worcs_repos$owner), function(x){
  owner_repos <- worcs_repos[worcs_repos$owner == x, , drop = FALSE]
  paste0("\n    - ", x, "  ",
         paste0("\n        + [", owner_repos$name, "](", owner_repos$url, ")  ", collapse = ""))
  }),
  collapse = "")

invited <- read.table("https://raw.githubusercontent.com/cjvanlissa/resume/master/invited.csv", sep = ",", header = TRUE)
invited <- invited[grepl("(worcs|workflow)", tolower(invited$Topic)), c("Institute", "Year", "City", "Country", "Date")]
invited <- paste0(apply(invited, 1, function(x){paste0("\n    - ", x[1], " (", x[2], ")  ")}), collapse = "")
```

As of `r Sys.Date()`, these are indicators of the adoption of `worcs` by users:

1. The preprint has been downloaded 702 times, since being published on 31-05-2020 (currently in press in [Data Science](https://content.iospress.com/journals/data-science/Pre-press/Pre-press))
1. The `worcs` R-package has been downloaded `r dl` times from CRAN, since being published on 18-05-2020
1. The GitHub project has been forked `r worcs_repo$network_count` times, watched `r worcs_repo$subscribers_count` times, and starred `r worcs_repo$watchers` times
1. The lead author has given invited lectures on WORCS at:`r invited`
1. WORCS is currently used in the following GitHub repositories (sorted by user):`r tmp`

## Contributing and Contact Information

We are always eager to receive user feedback and contributions to help us improve both the workflow and the software. Major contributions warrant coauthorship to the package. Please contact the lead author at [c.j.vanlissa@uu.nl](mailto:c.j.vanlissa@uu.nl), or:

  - [File a GitHub issue](https://github.com/cjvanlissa/worcs) for feedback, bug reports or feature requests
  - [Make a pull request](https://github.com/cjvanlissa/worcs/pulls) to contribute your code or prose

By participating in this project, you agree to abide by the [Contributor
Code of Conduct v2.0](https://www.contributor-covenant.org/). Contributions to the package must adhere to the [tidyverse style guide](https://style.tidyverse.org/). When contributing code, please add tests for that contribution to the `tests/testthat` folder, and ensure that these tests pass<!-- on [Travis-CI](https://travis-ci.org/github/cjvanlissa/worcs)-->.

## Acknowledgements

The worcs logo is inspired by the Open Science Badges by the Center for Open Science (CC-BY-4.0), and makes use of the gear, services, gears, preferences, settings icon, made by MD Badsha Meah from www.freeicons.io (CC-BY-3.0).